### 2017-05-16 experiment.Rmd

---
title: "Experiment"
date: "Last run: `r format(Sys.time(), '%F')`"
---

## Description

This survey was hosted on [Qualtrics](https://www.qualtrics.com) and participants were randomly sorted into one of eight variations (see [Design](design.html)).

Participants were recruited through [Amazon Mechanical Turk](https://www.mturk.com/mturk/welcome) and were paid $X for their response.

## Survey experiment {.survey}

### Consent

***Q1.1***: Consent form here…

*Q1.2*: I have read the above information, and I consent to take part in the study.

:   Single answer

    - Yes
    - No

---

### Crackdown and private giving

![](img/irc_logo.png "IRC logo"){.img-responsive .center}

\

***Q2.1***: The **International Rescue Committee** (IRC) focuses on {ISSUE AREA: humanitarian assistance for refugees | human rights for refugees} {MAIN MANIPULATION: and works in countries that have recently passed laws that harshly restrict nonprofit organizations}.

A substantial proportion of IRC’s funding comes from {FUNDING SOURCE: government | private donors}.

*Q2.2*: How likely would you be to donate to IRC?

:   Single answer

    - Extremely likely
    - Somewhat likely
    - Neither likely nor unlikely
    - Somewhat unlikely
    - Extremely unlikely

*Q2.3*: If you had an extra $100, how much would you be willing to donate to IRC right now? (*Total must add to 100*)

:   Numeric slider, options must sum to 100

    - Donate to IRC (0–100)
    - Keep for yourself (0–100)

*Q2.4*: Why would you donate that amount?

:   Text field

---

### Crackdown and corporate social responsibility (CSR)

![](img/google_irc_logo.png "Google and IRC logos"){.img-responsive .center}

\

***Q3.1***: In December 2016, Google held a Giving Weekend and raised $X and did YYY to support International Rescue Committee’s (IRC) efforts in {ISSUE AREA: providing humanitarian assistance for refugees | advocating human rights for refugees} {in countries where nonprofit organizations are harshly restricted}.

A substantial proportion of IRC’s funding comes from {FUNDING SOURCE: government | private donors}.

*Q3.2*: How favorable do you feel towards Google?

:   Single answer

    - Extremely positive
    - Somewhat positive
    - Neither positive nor negative
    - Somewhat negative
    - Extremely negative

*Q3.3*: How generous do you feel Google is?

:   Single answer

    - Extremely generous
    - Very generous
    - Moderately generous
    - Slightly generous
    - Not generous at all

*Q3.4*: Rate how strongly you feel Google is committed to philanthropy:

:   Single answer

    - Extremely committed
    - Very committed
    - Moderately committed
    - Slightly committed
    - Not committed at all

*Q3.5*: Where would you most like to see Google donate its money?

:   Single answer

    - Nonprofit organizations or foundations working mostly in California
    - Nonprofit organizations or foundations working throughout the United States
    - Nonprofit organizations or foundations working mostly in foreign countries

*Q3.6*: Why do you think Google should donate money to organizations working in X?

:   Text field

---

### Demographics

***Q4.1***: Almost done! This is the last page of the study.

*Q4.2*: How often do you donate to charity (either cash or in-kind)?

:   Single answer

    - Once a week
    - Once a month
    - Once every three months
    - Once every six months
    - Once a year
    - Once every few years
    - Never

*Q4.3*: Have you volunteered in the past 12 months?

:   Single answer

    - Yes
    - No

*Q4.4*: How closely do you follow politics in the news?

:   Single answer

    - Extremely closely
    - Very closely
    - Closely
    - Barely
    - Not at all

*Q4.5*: How would you describe your political views?

:   Single answer

    - Strong liberal
    - Liberal
    - Independent, leaning liberal
    - Independent
    - Independent, leaning conservative
    - Conservative
    - Very conservative

*Q4.6*: What is the highest degree or level of school you have completed?

:   Single answer

    - Less than high school
    - High school graduate
    - Some college
    - 2 year degree
    - 4 year degree
    - Graduate or professional degree
    - Doctorate

*Q4.7*: Aside from weddings and funerals, how often do you attend religious services?

:   Single answer

    - More than once a week
    - Once a week
    - Once or twice a month
    - A few times a year
    - Seldom
    - Never
    - Don't know

*Q4.8*: Please select blue from the following list:

:   Single answer

    - Red
    - Yellow
    - Blue
    - Green

*Q4.9*: What is your gender?

:   Single answer

    - Female
    - Male
    - Transgender
    - Other:
    - Prefer not to say

*Q4.10*: What is your total household income?

:   Single answer

    - Less than $10,000
    - $10,000 - $19,999
    - $20,000 - $29,999
    - $30,000 - $39,999
    - $40,000 - $49,999
    - $50,000 - $59,999
    - $60,000 - $69,999
    - $70,000 - $79,999
    - $80,000 - $89,999
    - $90,000 - $99,999
    - $100,000 - $149,999
    - More than $150,000

*Q4.11*: How old are you?

:   Single answer

    - Under 18
    - 18 - 24
    - 25 - 34
    - 35 - 44
    - 45 - 54
    - 55 - 64
    - 65 - 74
    - 75 - 84
    - 85 or older