### 2017-05-16 index.Rmd

---
title: "NGO Crackdowns and Philanthropy"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
---

\\(•◡•)/