### 2017-05-16 preregistration.Rmd

---
title: "Preregistration"
date: "Last run: `r format(Sys.time(), '%F')`"
---

## Center for Open Science preregistration

We have preregistered this study with the [Center for Open Science's](https://cos.io/) [Open Science Framework](https://osf.io/).

We are also participating in COS's [Preregistration Challenge](https://cos.io/our-services/prereg-more-information/), and will submit our final article to one of the following eligible journals:

- Voluntas
- International Interactions
- Journal of Experimental Political Science
- Political Science Research Methods
- Research and Politics

## Preregistration protocol

Our protocol is available at OSF and is included below.