### 2017-05-18 preregistration.Rmd
Italicize journal titles

---
title: "Preregistration"
date: "Last run: `r format(Sys.time(), '%F')`"
---

## Center for Open Science preregistration

We have preregistered this study with the [Center for Open Science's](https://cos.io/) [Open Science Framework](https://osf.io/).

We are also participating in COS's [Preregistration Challenge](https://cos.io/our-services/prereg-more-information/), and will submit our final article to one of the following eligible journals:

- *Voluntas*
- *International Interactions*
- *Journal of Experimental Political Science*
- *Political Science Research Methods*
- *Research and Politics*

## Preregistration protocol

Our protocol is available at OSF and is included below.