### 2018-02-09 01_clean-data.Rmd
Move stuff from older R files to Rmd

---
title: "Clean data"
author: "Andrew Heiss and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r clean-data}
library(tidyverse)
library(here)

results_file <- file.path(here(), "Data", "test_data.csv")

# Possible answers
favorability <- c("Very unfavorable", "Unfavorable", "Neutral", 
                  "Favorable", "Very favorable")
likelihood <- c("Extremely unlikely", "Somewhat unlikely", 
                "Neither likely nor unlikely", 
                "Somewhat likely", "Extremely likely")
frequency_charity <- c("Once a week", "Once a month", "Once every three months", 
                       "Once every six months", "Once a year", "Once every few years", 
                       "Never")
frequency_religion <- c("More than once a week", "Once a week", "Once or twice a month", 
                        "A few times a year", "Seldom", "Never", "Don't know")
income <- c("Less than $10,000", "$10,000 – $19,999", "$20,000 – $29,999", 
            "$30,000 – $39,999", "$40,000 – $49,999", "$50,000 – $59,999", 
            "$60,000 – $69,999", "$70,000 – $79,999", "$80,000 – $89,999", 
            "$90,000 – $99,999", "$100,000 – $149,999", "More than $150,000", 
            "Prefer not to say")
age <- c("Under 18", "18 – 24", "25 – 34", "35 – 44", "45 – 54", 
         "55 – 64", "65 – 74", "75 – 84", "85 or older")

# Qualtrics stores question information in the first 3 rows. All we care about
# are the column names in the first row.
results_meta <- read_csv(results_file, n_max = 2)

# The first 8 rows need to be skipped because the consent question has a bunch
# of line breaks and it messes with CSV line counting. The actual responses
# start on row 9. This doesn't entirely make sense, since read_csv can read the
# line break-filled row in just fine. Skipping said lines makes it choke. <shrug>
results_raw <- read_csv(results_file, skip = 8,
                        col_names = colnames(results_meta))

results <- results_raw %>%
  # Only select necessary columns
  select(id = ResponseId, duration = `Duration (in seconds)`,
         start_date = StartDate, end_date = EndDate,
         crackdown, issue, funding, starts_with("Q")) %>%
  rename(favor_humanitarian = Q2.1_1, favor_human_rights = Q2.1_2, 
         favor_development = Q2.1_3, donate_likely = Q2.3, 
         amount_donate = Q2.4_1, amount_keep = Q2.4_2, amount_why = Q2.5,
         give_charity = Q3.2, volunteer = Q3.3, political_knowledge = Q3.4,
         ideology = Q3.5, education = Q3.6, religiosity = Q3.7,
         gender = Q3.9, gender_other = Q3.9_4_TEXT, income = Q3.10, age = Q3.11) %>%
  # Clean up experimental condition columns
  mutate(issue = recode(issue, `human rights for refugees` = "Human rights",
                        `humanitarian assistance for refugees` = "Humanitarian assistance"),
         funding = recode(funding, `government donors` = "Government",
                          `individual, private donors` = "Private")) %>%
  mutate(crackdown = ifelse(str_detect(crackdown, "harshly restrict"), 
                            "Crackdown", "No crackdown"))
```