### 2018-02-09 02_results.Rmd

---
title: "Results"
author: "Andrew Heiss and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r skeleton, eval=FALSE}

# Check distribution of experimental conditions
results %>% count(crackdown, issue, funding)

# Average amount donated, by condition
# Individual frames
results %>% group_by(crackdown) %>% summarize(avg_donation = mean(amount_donate))
results %>% group_by(issue) %>% summarize(avg_donation = mean(amount_donate))
results %>% group_by(funding) %>% summarize(avg_donation = mean(amount_donate))
results %>% group_by(crackdown, funding) %>% summarize(avg_donation = mean(amount_donate))
results %>% group_by(crackdown, issue) %>% summarize(avg_donation = mean(amount_donate))
results %>% group_by(crackdown, issue, funding) %>% summarize(avg_donation = mean(amount_donate))

# Amount / propensity to donate based on predisposition towards nonprofits
# Amount / propensity to donate based on charitableness
# Amount / propensity to donate based on voluntarism
# Amount / propensity to donate based on political knowledge
# Amount / propensity to donate based on political ideology
# Amount / propensity to donate based on education
# Amount / propensity to donate based on religiosity
# Amount / propensity to donate based on gender
# Amount / propensity to donate based on income
# Amount / propensity to donate based on age

# Control for all those things

```

```{r stuff}
#   mutate(likelihood_bin = fct_recode(likelihood,
#                                      Unlikely = "Extremely unlikely",
#                                      Unlikely = "Somewhat unlikely",
#                                      Unlikely = "Neither likely nor unlikely",
#                                      Likely = "Somewhat likely",
#                                      Likely = "Extremely likely"))
#   
# plot(results$likelihood)
# plot(results$likelihood_bin)
```

```{r var-summary}
# results_summary <- results %>%
#   select(favorability, donation, keep,
#          crackdown, issue, funding)
```

```{r check-balance}
# balance <- results %>%
#   group_by(crackdown) %>%
#   summarise_at(vars(volunteered), mean)  # THIS IS WRONG THOUGH
```

```{r}
# wilcox.test(donation ~ crackdown, data = results) 
# t.test(donation ~ crackdown, data = results)
```

- Demographic balance across main crackdown condition

- Differences of these across conditions:

    - Mean likelihood
    - Mean likelihood collapsed to binary
    - Mean donation to NGO
    - Mean amount kept for person

- Modeling:

    - donation ~ frames (OLS)
    - likelihood binary  ~ frames (multinomial)
    - likelihood ~ frames (multinomial)
    - those models with controls