### 2018-06-25 experiment.Rmd
Move actual experiment to external file for reuse

---
title: "Experiment"
date: "Last run: `r format(Sys.time(), '%F')`"
---

# Description

This survey was hosted on [Qualtrics](https://www.qualtrics.com) and participants were randomly sorted into [one of eight conditions](preregistration.html#experimental_conditions).

Participants were recruited through [Amazon Mechanical Turk](https://www.mturk.com/mturk/welcome) and were paid \$0.75 for their response, which is equivalent to a \$9/hour wage:

```{r calculate-payment}
hourly_wage <- 9
survey_length_minutes <- 5

payment <- survey_length_minutes * (hourly_wage / 60)
payment
```

# Survey experiment {.survey}

```{r child = "text/experiment.md"}
```