### 2018-07-15 preregistration.Rmd
Move IRB text to separate file

---
title: "Preregistration, design, IRB"
date: "Last run: `r format(Sys.time(), '%F')`"
bibliography: manuscript/bibliography.bib
---

# Preregistration

We preregistered this study with the [Center for Open Science's](https://cos.io/) [Open Science Framework](https://osf.io/).

<p><a href="https://osf.io/dx973/register/565fb3678c5e4a66b5582f67" class="btn btn-primary btn-md btn-success" target="_blank">
<i class="fa fa-pencil-square" aria-hidden="true"></i>&nbsp;
OSF preregistration protocol &raquo;
</a></p>

We are participating in COS's [Preregistration Challenge](https://cos.io/our-services/prereg-more-information/), and will submit our final article to one of the following eligible journals:

- *Voluntas*
- *International Interactions*
- *Journal of Experimental Political Science*
- *Political Science Research Methods*
- *Research and Politics*

---

# Design

## Experimental conditions

| Condition            | 1 | 2 | 3 | 4 | 5 | 6 | 7 | 8 |
|----------------------|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|
| **Crackdown**        |   |   |   |   |   |   |   |   |
| Yes                  | • | • | • | • |   |   |   |   |
| No                   |   |   |   |   | • | • | • | • |
| *Expected donations* | + | + | + | + | ∅ | ∅ | ∅ | ∅ |
|                      |   |   |   |   |   |   |   |   |
| **Issue area**       |   |   |   |   |   |   |   |   |
| Humanitarian service | • | • |   |   | • | • |   |   |
| Human rights         |   |   | • | • |   |   | • | • |
| *Expected donations* | + | + | − | − | + | + | − | − |
|                      |   |   |   |   |   |   |   |   |
| **Type of funding**  |   |   |   |   |   |   |   |   |
| Government           | • |   | • |   | • |   | • |   |
| Private              |   | • |   | • |   | • |   | • |
| *Expected donations* | − | + | − | + | − | + | − | + |

## $f^2$ power analysis

The ultimate goal of our research project is to use Bayesian analysis is to test if the region of practical equivalence around the null value (i.e. the proposition that the average treatment effect of NGO crackdown is zero) excludes the posterior 95% HDI.

In this initial pilot study, we use standard frequentist methods to perform power analysis and estimate the desired sample size. We will use the effect size estimates from this study to inform the prior distribution of parameter values in future Bayesian-based experiments [@Kruschke:2015, 362].

```{r power-calculations}
library(pwr)
u_mine <- 1
# f2_mine <- cohen.ES(test = "f2", size = "small")$effect.size
f2_mine <- 0.02
sig_level_mine <- 0.05
power_mine = 0.90

survey_power <- pwr.f2.test(u = u_mine, v = NULL, f2 = f2_mine, 
                            sig.level = sig_level_mine, power = power_mine)

# n = u + v + 1
n_total <- survey_power$u + survey_power$v + 1
n_group <- n_total / (2 * 2 * 2)
```

Cohen's rules of thumb for $f^2$ power are 0.02, 0.15, and 0.35 for small, medium, and large effect sizes.

We are using a 2 × 2 × 2 design, manipulating the experimental condition, the issue area for the NGO, and the source of funding for the NGO. The formula for calculating the $f^2$-based sample size with [`pwr.f2.test`](https://www.rdocumentation.org/packages/pwr/versions/1.2-0/topics/pwr.f2.test) requires the following values:

- `u`: degrees of freedom for the numerator. This is a 2 × 2 × 2 design, so the degrees of freedom = $(2 - 1) \times (2 - 1) \times (2 - 1) = 1$
- `f2`: effect size. Here we use a small size, or `r f2_mine`.
- `sig.level`: significance level, or Type I error probability. Here we use `r sig_level_mine`.
- `power`: power of test, or 1 − Type II error probability. Here we use `r power_mine`.

Given these parameters, `pwr.f2.test` yields `v`, or the degrees of freedom for the denominator. The final sample size `n` is calculated as `u + v + 1`.

Thus, the total sample size is **`r round(n_total, 2)`**, with **`r round(n_group, 2)`** spread across 8 groups.

---

# Human subjects and IRB approval

This project has received IRB approval from the human subjects research committees at both BYU and Christopher Newport. We have included a generic version of our protocol below.

<p><a href="files/byu-irb-approval.pdf" class="btn btn-primary btn-md btn-success" target="_blank">
<i class="fa fa-check-circle" aria-hidden="true"></i>&nbsp;
BYU IRB approval
</a> 
&nbsp;
<a href="files/cnu-irb-approval.pdf" class="btn btn-primary btn-md btn-success" target="_blank">
<i class="fa fa-check-circle" aria-hidden="true"></i>&nbsp;
Christopher Newport IRB approval
</a></p>

\

```{r child = "text/irb.md"}
```

# References