### 2018-07-16 index.Rmd
Use safesite

---
title: "NGO Crackdowns and Philanthropy"
author: "Suparna Chaudhry and Andrew Heiss"
date: "Last run: `r format(Sys.time(), '%F')`"
site: safesite::safesite
---

\\(•◡•)/