### 2019-02-26 02_analysis.Rmd
Use RAW STAN for better analysis
No more triple interaction terms!

---
title: "Results"
author: "Andrew Heiss and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(rstan)
library(tidybayes)
library(broom)
library(ggstance)
library(patchwork)
library(pander)
library(scales)
library(janitor)
library(huxtable)
library(here)

source(here("lib", "graphics.R"))
source(here("lib", "pander_options.R"))
source(here("lib", "modeling.R"))

# Load data
results <- readRDS(here("data", "results_clean.rds"))
# qwraps2::lazyload_cache_dir("02_analysis_cache/html")

# Helpful functions
safe_mean <- function(x, ...) suppressWarnings(mean(as.numeric(x), ...))
safe_sd <- function(x, ...) suppressWarnings(sd(as.numeric(x), ...))

my_as_numeric <- function(x) {
  is_num <- suppressWarnings(!any(is.na(as.numeric(x))))

  if (is_num) {
    as.numeric(x)
  } else {
    x
  }
}

level_summary <- function(x, variable, df = results) {
  variable <- as.character(variable)
  if (is.factor(df[[variable]])) {
    x_levels <- levels(df[[variable]])
  } else {
    x_levels <- unique(df[[variable]])
  }

  factor_summary <- tibble(x) %>% 
    count(x) %>%
    mutate(pct = n / sum(n),
           fancy = paste0(x, " (", n, "; ", percent(pct), ")"),
           x = factor(x, levels = x_levels, ordered = TRUE))

  inline_summary <- factor_summary %>% arrange(x) %>% 
    pull(fancy) %>% paste(collapse = " | ")

  fig_bar <- ggplot(factor_summary, aes(x = x, y = n)) +
    geom_bar(stat = "identity", fill = "grey50") +
    theme_spark()

  file_name <- gsub("\\.| ", "_", tolower(variable))
  md_img <- paste0("![](output/figures/summary-table/", file_name, ")")

  ggsave(fig_bar, filename = here("output", "figures", "summary-table",
                                  paste0(file_name, ".pdf")), 
         width = 1, height = 0.2, device = cairo_pdf)
  ggsave(fig_bar, filename = here("output", "figures", "summary-table",
                                  paste0(file_name, ".png")), 
         width = 1, height = 0.2, type = "cairo", bg = "transparent", dpi = 300)

  output <- tibble(spark = md_img, summary = inline_summary)
  return(output)
}

numeric_summary <- function(x, variable) {
  inline_summary <- paste0("Median: ", round(median(x, na.rm = TRUE), 2),
                           " | Mean: ", round(mean(x, na.rm = TRUE), 2), 
                           " | Std. Dev.: ", round(sd(x, na.rm = TRUE), 2))

  fig_hist <- ggplot(as.data.frame(x), aes(x = x)) +
    geom_histogram(fill = "grey50", binwidth = 10) +
    theme_spark()

  file_name <- gsub("\\.| ", "_", tolower(variable))
  md_img <- paste0("![](output/figures/summary-table/", file_name, ")")

  ggsave(fig_hist, filename = here("output", "figures", "summary-table",
                                   paste0(file_name, ".pdf")), 
         width = 1, height = 0.2, device = cairo_pdf)
  ggsave(fig_hist, filename = here("output", "figures", "summary-table",
                                   paste0(file_name, ".png")), 
         width = 1, height = 0.2, type = "cairo", bg = "transparent", dpi = 300)

  output <- tibble(spark = md_img, summary = inline_summary)
  return(output)
}

my_summary <- function(x, variable, df) {
  x <- my_as_numeric(x)

  if (is.numeric(x)) {
    numeric_summary(x, variable)
  } else {
    level_summary(x, variable, df)
  }
}
```

# Overview of data

## Balance of experimental conditions

```{r check-conditions, results="asis"}
results %>% count(crackdown, issue, funding) %>% 
  rename(Crackdown = crackdown, Issue = issue, Funding = funding) %>% 
  janitor::adorn_totals(., where = "row") %T>% 
  pandoc.table() %>% 
  pandoc.table.return(caption = "Balance of experimental conditions {#tbl:experimental-conditions}") %>% 
  cat(file = here("output", "tables", "tbl-experimental-conditions.md"))
```

## Descriptive statistics table

```{r tbl-descriptive-statistics, results="asis", warning=FALSE}
vars_to_summarize <- tribble(
  ~variable, ~clean_name,
  "donate_likely", "Likelihood of donation",
  "donate_likely_bin", "Likelihood of donation (binary)",
  "amount_donate", "Amount hypothetically donated ($)",
  "gender", "Gender",
  "age", "Age",
  "income", "Income",
  "education", "Education",
  "religiosity", "Frequency of attending religious services",
  "ideology", "Political views",
  "political_knowledge", "Frequency of following public affairs",
  "give_charity", "Frequency of charitable donations",
  "volunteer", "Volunteered in past 12 months",
  "favor_humanitarian", "Prior favorability towards humanitarian NGOs",
  "favor_humanitarian_bin", "Prior favorability towards humanitarian NGOs (binary)",
  "favor_human_rights", "Prior favorability towards human rights NGOs",
  "favor_human_rights_bin", "Prior favorability towards human rights NGOs (binary)",
  "favor_development", "Prior favorability towards development NGOs",
  "favor_development_bin", "Prior favorability towards development NGOs (binary)",
  "check2", "Attention check 2"
)

results_summary_stats <- results %>% 
  select(one_of(vars_to_summarize$variable)) %>% 
  gather(variable, value) %>% 
  group_by(variable) %>% 
  nest() %>% 
  mutate(N = data %>% map_int(~ nrow(.)),
         summary = map2(.x = data, .y = variable, ~ my_summary(.x$value, .y, results))) %>% 
  left_join(vars_to_summarize, by = "variable") %>% 
  mutate(variable = factor(variable, levels = vars_to_summarize$variable, ordered = TRUE)) %>% 
  arrange(variable) %>% 
  select(-data, -variable) %>% 
  unnest(summary) %>% 
  select(Variable = clean_name, N, ` ` = spark, Details = summary)

results_summary_stats %>% 
  select(-N) %>% 
  pandoc.table.return(caption = "Descriptive statistics {#tbl:descriptive-stats}",
                      split.cell = 80, split.table = Inf) %T>% 
  cat(file = here("output", "tables", "tbl-descriptive-stats.md")) %>%
  cat()
```

## Average likelihood and amount donated across conditions

```{r tbl-results-conditions, results="asis"}
conditions_summary <- bind_rows(group_by(results, crackdown, issue, funding) %>% nest(),
                                group_by(results, crackdown, issue) %>% nest(),
                                group_by(results, crackdown) %>% nest(),
                                results %>% nest()) %>% 
  arrange(crackdown, issue, funding) %>% 
  mutate(summary = data %>%
           map(~ summarize(., pct_likely = table(donate_likely_bin)[["Likely"]] /
                             length(donate_likely_bin),
                           mean_donation = mean(amount_donate, na.rm = TRUE),
                           sd_donation = sd(amount_donate, na.rm = TRUE),
                           N = nrow(.)))) %>% 
  unnest(summary) %>% select(-data)

conditions_summary_clean <- conditions_summary %>% 
  mutate(funding = ifelse(is.na(funding) & !is.na(issue) , "*Total*", as.character(funding)),
         issue = ifelse(is.na(issue) & !is.na(crackdown), "*Total*", as.character(issue)),
         crackdown = ifelse(is.na(crackdown), "*Total*", as.character(crackdown))) %>% 
  group_by(crackdown) %>% 
  mutate(issue = replace(issue, duplicated(issue), NA)) %>% 
  ungroup() %>% 
  mutate(crackdown = replace(crackdown, duplicated(crackdown), NA)) %>% 
  mutate(pct_likely = percent(pct_likely)) %>% 
  rename(`Crackdown condition` = crackdown, `Issue condition` = issue,
         `Funding condition` = funding, `% likely to donate` = pct_likely,
         `Amount donated (mean)` = mean_donation, `Amount donated (sd)` = sd_donation)

conditions_summary_clean %T>% 
  pandoc.table() %>% 
  pandoc.table.return(caption = "Average likelihood and amount donated across experimental conditions {#tbl:avg-results}") %>% 
  cat(file = here("output", "tables", "tbl-avg-results.md"))
```

\

# Visualize important variables

## Likelihood of donation

```{r fig-likelihood-bars, fig.width=9, fig.height=4}
donate_summary <- results %>% 
  count(donate_likely) %>% 
  mutate(perc = n / sum(n)) %>% 
  mutate(highlight = ifelse(donate_likely %in% c("Extremely likely", "Somewhat likely"), TRUE, FALSE))

plot_donate_summary <- ggplot(donate_summary, aes(x = n, y = donate_likely, 
                                                  fill = highlight)) +
  geom_barh(stat = "identity") +
  scale_x_continuous(sec.axis = sec_axis(~ . / sum(donate_summary$n),
                                         labels = percent_format(accuracy = 1))) +
  scale_fill_manual(values = ngo_cols("green", "blue", name = FALSE), guide = FALSE) +
  labs(x = NULL, y = NULL) +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

plot_donate_summary %T>% 
  print() %T>%
  ggsave(., filename = here("output", "figures", "donate_summary.pdf"),
         width = 9, height = 4, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "figures", "donate_summary.png"),
         width = 9, height = 4, units = "in", type = "cairo", dpi = 300)
```

## Amount donated

```{r fig-amount-bars, fig.width=9, fig.height=2.75}
plot_amount_summary <- ggplot(results, aes(x = amount_donate)) +
  geom_histogram(bins = 20, fill = ngo_cols("blue")) +
  scale_x_continuous(labels = dollar) +
  labs(x = NULL, y = NULL) +
  theme_ngos() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())

plot_amount_summary %T>% 
  print() %T>%
  ggsave(., filename = here("output", "figures", "amount_summary.pdf"),
         width = 9, height = 2.75, units = "in", device = cairo_pdf) %>% 
  ggsave(., filename = here("output", "figures", "amount_summary.png"),
         width = 9, height = 2.75, units = "in", type = "cairo", dpi = 300)
```

\

```{r all-datasets-nested}
all_models <- tribble(
  ~title, ~df,
  "Crackdown", results,
  # Ha, cheat here by renaming the issue and funding columns to crackdown. 
  # All other tests are based on the crackdown column except these two, so
  # rather than build all sorts of convoluted arguments and functions, we just
  # rename these as issue and funding
  "Issue", mutate(results, crackdown = issue),
  "Funding", mutate(results, crackdown = funding),
  # Create all the nested conditions
  "Human rights | Crackdown", filter(results, issue == "Human rights"),
  "Humanitarian assistance | Crackdown", filter(results, issue != "Human rights"),
  "Government | Crackdown", filter(results, funding == "Government"),
  "Private | Crackdown", filter(results, funding == "Private"),
  "Human rights | Government | Crackdown", filter(results, issue == "Human rights", funding == "Government"),
  "Human rights | Private | Crackdown", filter(results, issue == "Human rights", funding == "Private"),
  "Humanitarian assistance | Government | Crackdown", filter(results, issue != "Human rights", funding == "Government"),
  "Humanitarian assistance | Private | Crackdown", filter(results, issue != "Human rights", funding == "Private")
) 
```

# Treatment effects: Likelihood of donation

This time around, we're not using interactionful regression models to calculate all these differences in groups. Instead we use real live Stan code to estimate the differences in group means and proportions!

## Priors and models

We estimate the proportion of people responding that they'd be likely to donate to the organization with a binomial distribution, with a prior $\theta$ distribution of $\text{Beta}(5, 5)$. We build the following model in Stan:

$$
\begin{aligned}
n_{\text{group 1, group 2}} &\sim \text{Binomial}(n_{\text{total in group}}, \theta_{\text{group}}) &\text{[likelihood]}\\
\text{Difference} &= n_{\text{group 2}} - n_{\text{group 1}} &\text{[difference in proportions]} \\
n &: \text{Number likely to donate} \\
\\
\theta_{\text{group 1, group 2}} &\sim \text{Beta}(5, 5) &\text{[prior prob. of being likely to donate]}
\end{aligned}
$$

```{r priors-likely, fig.width=2.75, fig.height=1.5, fig.align="center"}
likely_theta <- ggplot(data = tibble(x = c(0, 1)), aes(x = x)) +
  geom_area(stat = "function", fun = dbeta, args = list(shape1 = 5, shape2 = 5), 
            fill = "grey80", color = "black") +
  labs(x = expression(Probability ~ of ~ being ~ likely ~ to ~ donate ~ (theta)), y = "Density") +
  annotate(geom = "label", x = 0.5, y = 1, label = "Beta(5, 5)", size = pts(9)) +
  theme_ngos(base_size = 9, density = TRUE)
likely_theta
```

## Differences

```{r run-likely-models, results="hide", cache=TRUE}
# This takes forever because of C++ compiling and actual MCMC sampling, so this
# chunk is cached

# Compile model
donate_likely <- stan_model(here("lib", "donate_likely.stan"))

# Sample from the compiled model
all_models_likely_run <- all_models %>% 
  # Simplify data for sending to Stan
  mutate(stan_data_counts = df %>% map(~ {
    df_counts <- .x %>% 
      count(crackdown, donate_likely_bin) %>% 
      group_by(crackdown) %>% 
      mutate(total = sum(n)) %>% 
      filter(donate_likely_bin == "Likely")

    return(list(
      n_total_1 = df_counts$total[1],
      n_total_2 = df_counts$total[2],
      n_likely_1 = df_counts$n[1],
      n_likely_2 = df_counts$n[2]
    ))
  })) %>% 
  # Run the actual model on the simplified data
  mutate(model = stan_data_counts %>% map(~ {
    sampling(donate_likely, data = .x, control = list(max_treedepth = 15),
             chains = CHAINS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED)
  }))

# Extract stuff
all_models_likely <- all_models_likely_run %>% 
  # Extract posterior chains
  mutate(posterior_chains_long = map(model, ~ {
      .x %>% gather_draws(theta_1, theta_2, theta_diff, pct_change)
  })) %>% 
  mutate(posterior_chains_wide = map(model, ~ {
      .x %>% spread_draws(theta_1, theta_2, theta_diff, pct_change)
  })) %>% 
  # Get HDI medians in tidy form
  mutate(tidy = posterior_chains_long %>% map(~ {
    .x %>% median_hdci() %>% to_broom_names()
  }))
```

```{r plot-likely-diffs, warning=FALSE, fig.width=6, fig.height=8}
tidied_diffs_likely <- all_models_likely %>% 
  unnest(posterior_chains_long) %>% 
  filter(.variable == "theta_diff") %>% 
  mutate(category = case_when(
    str_count(title, "\\|") == 0 ~ "Level 1",
    str_count(title, "\\|") == 1 ~ "Level 2",
    str_count(title, "\\|") == 2 ~ "Level 3"
  )) %>% 
  mutate(title = str_remove(title, " \\| Crackdown"))

level1_likely <- tidied_diffs_likely %>% 
  filter(category == "Level 1") %>%
  mutate(title = recode(title, 
                        Crackdown = "Crackdown −\nNo crackdown",
                        Issue = "Humanitarian\nassistance −\nHuman rights",
                        Funding = "Private −\nGovernment\nfunding")) %>% 
  mutate(title = fct_inorder(title))

plot_diff_likely_a <- ggplot(level1_likely, aes(x = .value, y = fct_rev(title), fill = title)) +
  geom_halfeyeh(.width = c(0.8, 0.95)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = ngo_cols(c("blue", "red", "orange"), name = FALSE), guide = FALSE) +
  labs(x = "Median difference in donation likelihood", y = NULL, tag = "A") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

level2_likely <- tidied_diffs_likely %>% 
  filter(category == "Level 2") %>%
  mutate(condition = case_when(
    title %in% c("Human rights", "Humanitarian assistance") ~ "Issue",
    title %in% c("Government", "Private") ~ "Funding"
  )) %>% 
  mutate(facet_title = case_when(
    title %in% c("Human rights", "Humanitarian assistance") ~ paste(title, "issues"),
    title %in% c("Government", "Private") ~ paste(title, "funding")
  )) %>% 
  mutate(title = case_when(
    title %in% c("Humanitarian assistance", "Private") ~ "",
    TRUE ~ "Crackdown −\nNo crackdown")
  )

plot_diff_likely_b <- ggplot(filter(level2_likely, condition == "Issue"), 
                             aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95), fill = ngo_cols("red", name = FALSE)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL, tag = "B") +
  facet_wrap(~ facet_title, scales = "free_y") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

plot_diff_likely_c <- ggplot(filter(level2_likely, condition == "Funding"), 
                             aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95), fill = ngo_cols("orange", name = FALSE)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Median difference in donation likelihood", y = NULL, tag = "C") +
  facet_wrap(~ facet_title, scales = "free_y") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

level3_likely <- tidied_diffs_likely %>% 
  filter(category == "Level 3") %>%
  separate(title, c("issue", "funding"), sep = " \\| ") %>%
  mutate(issue = paste(issue, "issues"),
         funding = paste(funding, "funding")) %>% 
  mutate(facet_title = paste0(issue, "\n", funding)) %>% 
  mutate(title = case_when(
    funding == "Private funding" ~ "",
    TRUE ~ "Crackdown −\nNo crackdown")
  )

plot_diff_likely_d <- ggplot(level3_likely, aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95), fill = ngo_cols("green", name = FALSE)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Median difference in donation likelihood", y = NULL, tag = "D") +
  facet_wrap(~ facet_title, scales = "free_y") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

plot_diffs_likely_all <- plot_diff_likely_a / 
  (plot_diff_likely_b / plot_diff_likely_c) / 
  plot_diff_likely_d

plot_diffs_likely_all

plot_diffs_likely_all %T>% 
  ggsave(., filename = here("output", "figures", "likely-diffs.pdf"),
         width = 6, height = 8, units = "in", device = cairo_pdf) %>%
  ggsave(., filename = here("output", "figures", "likely-diffs.png"),
         width = 6, height = 8, units = "in", type = "cairo", dpi = 300)
```

```{r tbl-likely-diffs-level1, warning=FALSE, message=FALSE, results="asis"}
tbl_likely_tidy <- all_models_likely %>% 
  unnest(tidy) %>% 
  filter(term %in% c("theta_diff", "pct_change")) %>% 
  select(title, term, estimate) %>% 
  spread(term, estimate) %>% 
  mutate(category = case_when(
    str_count(title, "\\|") == 0 ~ "Level 1",
    str_count(title, "\\|") == 1 ~ "Level 2",
    str_count(title, "\\|") == 2 ~ "Level 3"
  ))

tbl_likely_probs <- all_models_likely %>% 
  unnest(posterior_chains_long) %>% 
  filter(.variable == "theta_diff") %>% 
  group_by(title) %>% 
  summarize(p.greater0 = mean(.value > 0),
            p.less0 = mean(.value < 0),
            p.diff.not0 = ifelse(median(.value) > 0, p.greater0, p.less0)) %>% 
  ungroup()

tbl_likely_tidy %>% 
  left_join(tbl_likely_probs, by = "title") %>% 
  filter(category == "Level 1") %>%
  mutate(title = recode(title, 
                        Crackdown = "Crackdown −\nNo crackdown",
                        Issue = "Humanitarian\nassistance −\nHuman rights",
                        Funding = "Private −\nGovernment\nfunding")) %>% 
  mutate(title = fct_inorder(title)) %>% 
  select(`Frame` = title, `∆` = theta_diff, `%∆` = pct_change,
         `p(∆ > 0)` = p.greater0, `p(∆ < 0)` = p.less0, `p(∆ ≠ 0)` = p.diff.not0) %>% 
  pandoc.table.return(caption = 'Difference between likelihood of donation across single conditions; values represent posterior medians {#tbl:likely-diffs-1}') %T>%
  cat() %>% 
  cat(file = here("output", "tables", "tbl-likely-diffs-1.md"))
```

```{r tbl-likely-diffs-level2, warning=FALSE, message=FALSE, results="asis"}
tbl_likely_tidy %>% 
  left_join(tbl_likely_probs, by = "title") %>% 
  filter(category == "Level 2") %>%
  mutate(title_clean = str_remove(title, " \\| Crackdown")) %>% 
  mutate(title_clean = case_when(
    title_clean %in% c("Human rights", "Humanitarian assistance") ~ paste(title_clean, "issues"),
    title_clean %in% c("Government", "Private") ~ paste(title_clean, "funding")
  )) %>% 
  mutate(title_clean = factor(title_clean, 
                              levels = c("Human rights issues", 
                                         "Humanitarian assistance issues", 
                                         "Government funding", "Private funding"))) %>% 
  arrange(title_clean) %>% 
  select(Frame = title_clean, `∆` = theta_diff, `%∆` = pct_change,
         `p(∆ > 0)` = p.greater0, `p(∆ < 0)` = p.less0, `p(∆ ≠ 0)` = p.diff.not0) %>% 
  pandoc.table.return(caption = 'Difference between likelihood of donation in "crackdown" (treatment) and "no crackdown" (control) conditions; values represent posterior medians {#tbl:likely-diffs-2}') %T>%
  cat() %>% 
  cat(file = here("output", "tables", "tbl-likely-diffs-2.md"))
```

```{r tbl-likely-diffs-level3, warning=FALSE, message=FALSE, results="asis"}
tbl_likely_tidy %>% 
  left_join(tbl_likely_probs, by = "title") %>% 
  filter(category == "Level 3") %>%
  mutate(title_clean = str_remove(title, " \\| Crackdown")) %>% 
  separate(title_clean, c("Issue", "Funding"), sep = " \\| ") %>% 
  mutate(Issue = factor(Issue, levels = c("Human rights", "Humanitarian assistance")),
         Funding = factor(Funding, levels = c("Government", "Private"))) %>% 
  arrange(Issue, Funding) %>% 
  select(`Issue frame` = Issue, `Funding frame` = Funding, `∆` = theta_diff, `%∆` = pct_change,
         `p(∆ > 0)` = p.greater0, `p(∆ < 0)` = p.less0, `p(∆ ≠ 0)` = p.diff.not0) %>% 
  pandoc.table.return(caption = 'Difference between likelihood of donation in "crackdown" (treatment) and "no crackdown" (control) conditions; values represent posterior medians {#tbl:likely-diffs-3}') %T>%
  cat() %>% 
  cat(file = here("output", "tables", "tbl-likely-diffs-3.md"))
```

\

# Treatment effects: Amount donated

## Priors and models

Following [John Kruschke's "Bayesian Estimation Supersedes the t-test (BEST)" procedure](http://www.indiana.edu/~kruschke/BEST/BEST.pdf), we estimate means for each group with a t-distribution. We use the following priors for the distribution parameters:

$$
\begin{aligned}
x_{\text{group 1, group 2}} &\sim \text{Student } t(\nu, \mu, \sigma) &\text{[likelihood]}\\
\text{Difference} &= x_{\text{group 2}} - x_{\text{group 1}} &\text{[difference in means]} \\
x &: \text{Mean amount donated} \\
\\
\nu &\sim \text{Exponential}(1 / 29) &\text{[prior normality]} \\
\mu_{\text{group 1, group 2}} &\sim \mathcal{N}(\bar{x}_{\text{group 1, group 2}}, 10) &\text{[prior donation mean per group]}\\
\sigma_{\text{group 1, group 2}} &\sim \text{Cauchy}(0, 1)&\text{[prior donation sd per group]}
\end{aligned}
$$

```{r priors-amounts, fig.width=8, fig.height=1.5}
amount_nu <- ggplot(data = tibble(x = c(0, 200)), aes(x = x)) +
  geom_area(stat = "function", fun = dexp, args = list(rate = 1/29), 
            fill = "grey80", color = "black") +
  labs(x = expression(Normality ~ parameter ~ (nu)), y = "Density") +
  annotate(geom = "label", x = 100, y = 0.009, 
           label = "Exponential(1/29)", size = pts(9)) +
  theme_ngos(base_size = 9, density = TRUE)

amount_mu <- ggplot(data = tibble(x = c(0, 100)), aes(x = x)) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 50, sd = 10),
            fill = "grey80", color = "black") +
  scale_x_continuous(breaks = c(seq(0, 100, 25)),
                     labels = c("−$50", "−$25", "Group average", "+$25", "+$50")) +
  annotate(geom = "label", x = 50, y = 0.01, label = "N(bar(x), 10)", 
           parse = TRUE, size = pts(9)) +
  labs(x = expression(Average ~ donated ~ (mu)), y = NULL) +
  theme_ngos(base_size = 9, density = TRUE)

amount_sigma <- ggplot(data = tibble(x = c(0, 10)), aes(x = x)) +
  geom_area(stat = "function", fun = dcauchy, args = list(location = 0, scale = 1),
            fill = "grey80", color = "black") +
  scale_x_continuous(labels = dollar) +
  annotate(geom = "label", x = 5, y = 0.08, label = "Cauchy(0, 1)", size = pts(9)) +
  labs(x = expression(SD ~ donated ~ (sigma)), y = NULL) +
  theme_ngos(base_size = 9, density = TRUE)

amount_nu + amount_mu + amount_sigma
```

## Differences

```{r run-amount-models, results="hide", cache=TRUE}
# This takes forever because of C++ compiling and actual MCMC sampling, so this
# chunk is cached

# Compile model
amount_donated_best <- stan_model(here("lib", "amount_donated_best.stan"))

# Sample from the compiled model
all_models_amount_run <- all_models %>% 
  # Simplify data for sending to Stan
  mutate(stan_data = df %>% map(~ {
    select(.x, amount = amount_donate, group = crackdown)
  })) %>% 
  # Run the actual model on the simplified data
  mutate(model = stan_data %>% map(~ {
    sampling(amount_donated_best, data = compose_data(.x),
             chains = CHAINS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED)
  }))

# Extract stuff
all_models_amount <- all_models_amount_run %>% 
  # Extract posterior chains
  mutate(posterior_chains_long = map2(model, stan_data, ~ {
    .x %>% 
      recover_types(.y) %>% 
      gather_draws(mu[group], sigma[group], mu_diff, pct_change, cohen_d, cles, nu, log10nu)
  })) %>% 
  mutate(posterior_chains_wide = map2(model, stan_data, ~ {
    .x %>% 
      recover_types(.y) %>% 
      spread_draws(mu[group], sigma[group], mu_diff, pct_change, cohen_d, cles, nu, log10nu)
  })) %>% 
  # Get HDI medians in tidy form
  mutate(tidy = posterior_chains_long %>% map(~ {
    .x %>% median_hdci() %>% to_broom_names()
  }))
```

```{r plot-amount-diffs, warning=FALSE, fig.width=6, fig.height=8}
# This will generate a warning because unnest() puts the crackdown values into
# one column and we cheated by renaming issue and funding as crackdown, so it
# gets mad when combining the factors in crackdown with those in issue and
# funding. But because we only care about the diffs, we don't actually need the
# category labels here, so it's all good.
tidied_diffs_amount <- all_models_amount %>% 
  unnest(posterior_chains_long) %>% 
  filter(.variable == "mu_diff") %>% 
  mutate(category = case_when(
    str_count(title, "\\|") == 0 ~ "Level 1",
    str_count(title, "\\|") == 1 ~ "Level 2",
    str_count(title, "\\|") == 2 ~ "Level 3"
  )) %>% 
  mutate(title = str_remove(title, " \\| Crackdown"))

level1_amount <- tidied_diffs_amount %>% 
  filter(category == "Level 1") %>%
  mutate(title = recode(title, 
                        Crackdown = "Crackdown −\nNo crackdown",
                        Issue = "Humanitarian\nassistance −\nHuman rights",
                        Funding = "Private −\nGovernment\nfunding")) %>% 
  mutate(title = fct_inorder(title))

plot_diff_amount_a <- ggplot(level1_amount, aes(x = .value, y = fct_rev(title), fill = title)) +
  geom_vline(xintercept = 0) +
  geom_halfeyeh(.width = c(0.8, 0.95)) +
  scale_x_continuous(labels = dollar) +
  scale_fill_manual(values = ngo_cols(c("blue", "red", "orange"), name = FALSE), guide = FALSE) +
  labs(x = "Median difference in amount donated", y = NULL, tag = "A") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

level2_amount <- tidied_diffs_amount %>% 
  filter(category == "Level 2") %>%
  mutate(condition = case_when(
    title %in% c("Human rights", "Humanitarian assistance") ~ "Issue",
    title %in% c("Government", "Private") ~ "Funding"
  )) %>% 
  mutate(facet_title = case_when(
    title %in% c("Human rights", "Humanitarian assistance") ~ paste(title, "issues"),
    title %in% c("Government", "Private") ~ paste(title, "funding")
  )) %>% 
  mutate(title = case_when(
    title %in% c("Humanitarian assistance", "Private") ~ "",
    TRUE ~ "Crackdown −\nNo crackdown")
  )

plot_diff_amount_b <- ggplot(filter(level2_amount, condition == "Issue"), 
                             aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95), fill = ngo_cols("red", name = FALSE)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = dollar) +
  labs(x = NULL, y = NULL, tag = "B") +
  facet_wrap(~ facet_title, scales = "free_y") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

plot_diff_amount_c <- ggplot(filter(level2_amount, condition == "Funding"), 
                             aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95), fill = ngo_cols("orange", name = FALSE)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = dollar) +
  labs(x = "Median difference in amount donated", y = NULL, tag = "C") +
  facet_wrap(~ facet_title, scales = "free_y") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

level3_amount <- tidied_diffs_amount %>% 
  filter(category == "Level 3") %>%
  separate(title, c("issue", "funding"), sep = " \\| ") %>%
  mutate(issue = paste(issue, "issues"),
         funding = paste(funding, "funding")) %>% 
  mutate(facet_title = paste0(issue, "\n", funding)) %>% 
  mutate(title = case_when(
    funding == "Private funding" ~ "",
    TRUE ~ "Crackdown −\nNo crackdown")
  )

plot_diff_amount_d <- ggplot(level3_amount, aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95), fill = ngo_cols("green", name = FALSE)) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(labels = dollar) +
  labs(x = "Median difference in amount donated", y = NULL, tag = "D",
       caption = "Point shows posterior median; thick black lines show 80% credible interval; thin black lines show 95% credible interval") +
  facet_wrap(~ facet_title, scales = "free_y") +
  theme_ngos() +
  theme(panel.grid.major.y = element_blank())

plot_diffs_amounts_all <- plot_diff_amount_a / 
  (plot_diff_amount_b / plot_diff_amount_c) / 
  plot_diff_amount_d

plot_diffs_amounts_all

plot_diffs_amounts_all %T>% 
  ggsave(., filename = here("output", "figures", "amount-diffs.pdf"),
         width = 6, height = 8, units = "in", device = cairo_pdf) %>%
  ggsave(., filename = here("output", "figures", "amount-diffs.png"),
         width = 6, height = 8, units = "in", type = "cairo", dpi = 300)
```

```{r tbl-amount-diffs-level1, warning=FALSE, message=FALSE, results="asis"}
tbl_amounts_tidy <- all_models_amount %>% 
  unnest(tidy) %>% 
  filter(term %in% c("mu_diff", "pct_change")) %>% 
  select(title, term, estimate) %>% 
  spread(term, estimate) %>% 
  mutate(category = case_when(
    str_count(title, "\\|") == 0 ~ "Level 1",
    str_count(title, "\\|") == 1 ~ "Level 2",
    str_count(title, "\\|") == 2 ~ "Level 3"
  ))

tbl_amounts_probs <- all_models_amount %>% 
  unnest(posterior_chains_long) %>% 
  filter(.variable == "mu_diff") %>% 
  group_by(title) %>% 
  summarize(p.greater0 = mean(.value > 0),
            p.less0 = mean(.value < 0),
            p.diff.not0 = ifelse(median(.value) > 0, p.greater0, p.less0)) %>% 
  ungroup()

tbl_amounts_tidy %>% 
  left_join(tbl_amounts_probs, by = "title") %>% 
  filter(category == "Level 1") %>%
  mutate(title = recode(title, 
                        Crackdown = "Crackdown −\nNo crackdown",
                        Issue = "Humanitarian\nassistance −\nHuman rights",
                        Funding = "Private −\nGovernment\nfunding")) %>% 
  mutate(title = fct_inorder(title)) %>% 
  select(`Frame` = title, `∆` = mu_diff, `%∆` = pct_change,
         `p(∆ > 0)` = p.greater0, `p(∆ < 0)` = p.less0, `p(∆ ≠ 0)` = p.diff.not0) %>% 
  pandoc.table.return(caption = 'Difference between amount donated across single conditions; values represent posterior medians {#tbl:amount-diffs-1}') %T>%
  cat() %>% 
  cat(file = here("output", "tables", "tbl-amount-diffs-1.md"))
```

```{r tbl-amount-diffs-level2, warning=FALSE, message=FALSE, results="asis"}
tbl_amounts_tidy %>% 
  left_join(tbl_amounts_probs, by = "title") %>% 
  filter(category == "Level 2") %>%
  mutate(title_clean = str_remove(title, " \\| Crackdown")) %>% 
  mutate(title_clean = case_when(
    title_clean %in% c("Human rights", "Humanitarian assistance") ~ paste(title_clean, "issues"),
    title_clean %in% c("Government", "Private") ~ paste(title_clean, "funding")
  )) %>% 
  mutate(title_clean = factor(title_clean, 
                              levels = c("Human rights issues", 
                                         "Humanitarian assistance issues", 
                                         "Government funding", "Private funding"))) %>% 
  arrange(title_clean) %>% 
  select(Frame = title_clean, `∆` = mu_diff, `%∆` = pct_change,
         `p(∆ > 0)` = p.greater0, `p(∆ < 0)` = p.less0, `p(∆ ≠ 0)` = p.diff.not0) %>% 
  pandoc.table.return(caption = 'Difference between amount donated in "crackdown" (treatment) and "no crackdown" (control) conditions; values represent posterior medians {#tbl:amount-diffs-2}') %T>%
  cat() %>% 
  cat(file = here("output", "tables", "tbl-amount-diffs-2.md"))
```

```{r tbl-amount-diffs-level3, warning=FALSE, message=FALSE, results="asis"}
tbl_amounts_tidy %>% 
  left_join(tbl_amounts_probs, by = "title") %>% 
  filter(category == "Level 3") %>%
  mutate(title_clean = str_remove(title, " \\| Crackdown")) %>% 
  separate(title_clean, c("Issue", "Funding"), sep = " \\| ") %>% 
  mutate(Issue = factor(Issue, levels = c("Human rights", "Humanitarian assistance")),
         Funding = factor(Funding, levels = c("Government", "Private"))) %>% 
  arrange(Issue, Funding) %>% 
  select(`Issue frame` = Issue, `Funding frame` = Funding, `∆` = mu_diff, `%∆` = pct_change,
         `p(∆ > 0)` = p.greater0, `p(∆ < 0)` = p.less0, `p(∆ ≠ 0)` = p.diff.not0) %>% 
  pandoc.table.return(caption = 'Difference between amount donated in "crackdown" (treatment) and "no crackdown" (control) conditions; values represent posterior medians {#tbl:amount-diffs-3}') %T>%
  cat() %>% 
  cat(file = here("output", "tables", "tbl-amount-diffs-3.md"))
```

## Effect size

```{r amount-effect-size, warning=FALSE}
# Effect size
# (μ₁ - μ₂) / sqrt( (σ₁² + σ₂²) / 2)
tidied_eff_size_amount <- all_models_amount %>% 
  unnest(posterior_chains_long) %>% 
  filter(.variable == "cohen_d") %>% 
  mutate(category = case_when(
    str_count(title, "\\|") == 0 ~ "Level 1",
    str_count(title, "\\|") == 1 ~ "Level 2",
    str_count(title, "\\|") == 2 ~ "Level 3"
  )) %>% 
  mutate(title = str_remove(title, " \\| Crackdown")) %>% 
  mutate(title = fct_inorder(title))

effect_sizes <- tribble(
  ~size, ~x_end,
  "Small", 0.2,
  "Medium", 0.5,
  "Large", 0.8
) %>% 
  mutate(x_start = -x_end)

ggplot(tidied_eff_size_amount, aes(x = .value, y = fct_rev(title))) +
  geom_halfeyeh(.width = c(0.8, 0.95)) +
  geom_vline(xintercept = 0, size = 1) +
  geom_vline(xintercept = c(0.2, -0.2), linetype = "dotted", size = 0.5) +
  geom_vline(xintercept = c(0.5, -0.5), linetype = "dotted", size = 0.5) +
  geom_vline(xintercept = c(0.5, -0.5), linetype = "dotted", size = 0.5) +
  coord_cartesian(xlim = c(-0.6, 0.6)) +
  labs(x = "Cohen's d (effect size)", y = NULL) +
  theme_ngos()
```

\

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>