### 2019-03-08 analysis/01_clean-data.Rmd
HUGE reorganization (go to 0290370 for previous history)
Switched this to build as a package using rrtools, with pre-built Stan binaries through rstantools. As a result, everything got shuffled around, and GitHub doesn't use --follow in their logs, so the history is hidden. Going to 02903709c87b79e8a945dd31fafd2484cbf2bc7b and going back in history from there will let you see past changes to these files.

In general, here's what happened:

- All analysis files in the project root (Rmd files that generated the website) are now in /analysis/*
- R functions that lived in lib/* are now standalone

---
title: "Clean data"
author: "Andrew Heiss and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Load and clean data

```{r clean-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(crackdownsphilanthropy)
library(csvy)
library(here)

# Load data
# results_file <- file.path(here(), "data", "test_data.csv")
results_file <- here("data", "private_data", 
                     "NGO Crackdowns and Philanthropy, March 2018 pilot_March 23, 2018_20.01.csv")

# Possible answers
favorability <- c("Very unfavorable", "Unfavorable", "Neutral", 
                  "Favorable", "Very favorable")

likelihood <- c("Extremely unlikely", "Somewhat unlikely", 
                "Neither likely nor unlikely", 
                "Somewhat likely", "Extremely likely")

frequency_charity <- c("Once a week", "Once a month", "Once every three months", 
                       "Once every six months", "Once a year", "Once every few years", 
                       "Never")

frequency_public_affairs <- c("Most of the time", "Some of the time", 
                              "Only now and then", "Hardly at all")

levels_ideology <- c("Strong liberal", "Liberal", "Independent, leaning liberal", 
                     "Independent", "Independent, leaning conservative", 
                     "Conservative", "Very conservative")

levels_education <- c("Less than high school", "High school graduate", 
                      "Some college", "2 year degree", "4 year degree", 
                      "Graduate or professional degree", "Doctorate")

frequency_religion <- c("More than once a week", "Once a week", "Once or twice a month", 
                        "A few times a year", "Seldom", "Never", "Don't know")

levels_gender <- c("Female", "Male", "Transgender", "Other", "Prefer not to say")

levels_income <- c("Less than $10,000", "$10,000 – $19,999", "$20,000 – $29,999", 
                   "$30,000 – $39,999", "$40,000 – $49,999", "$50,000 – $59,999", 
                   "$60,000 – $69,999", "$70,000 – $79,999", "$80,000 – $89,999", 
                   "$90,000 – $99,999", "$100,000 – $149,999", "More than $150,000", 
                   "Prefer not to say")

levels_age <- c("Under 18", "18 – 24", "25 – 34", "35 – 44", "45 – 54", 
                "55 – 64", "65 – 74", "75 – 84", "85 or older")

attention1_answer <- c("Green", "Yellow")
attention2_answer <- "Blue"
levels_attention <- c("Correct", "Incorrect")

# Qualtrics stores question information in the first 3 rows. All we care about
# are the column names in the first row.
results_meta <- read_csv(results_file, n_max = 2)

# The first 10 rows need to be skipped because the consent question has a bunch
# of line breaks and it messes with CSV line counting. The actual responses
# start on row 11. This doesn't entirely make sense, since read_csv can read the
# line break-filled row in just fine. Skipping said lines makes it choke. <shrug>
results_raw <- read_csv(results_file, skip = 16,
                        col_names = colnames(results_meta))

# Confirmation codes to exclude. These are people who failed the attention
# checks or who took the survey outside of MTurk.
#
# The code for determining these is in a script that is excluded from the main
# repository because of privacy issues (it deals directly with MTurk worker
# IDs), called "data/private_data/approve_mturkers.R". It outputs this text list
# of codes.
codes_to_exclude <- read_csv(here("data", "derived_data", "codes_to_exclude.csv"))

# Clean everything
results_check_attention <- results_raw %>%
  # Only select necessary columns
  select(id = ResponseId, confirmation_code = mTurkCode, duration = `Duration (in seconds)`,
         start_date = StartDate, end_date = EndDate,
         crackdown, issue, funding, starts_with("Q")) %>%
  rename(favor_humanitarian = Q2.1_1, favor_human_rights = Q2.1_2, 
         favor_development = Q2.1_3, donate_likely = Q2.3, 
         amount_donate = Q2.4_1, amount_keep = Q2.4_2, amount_why = Q2.5,
         give_charity = Q3.2, volunteer = Q3.3, political_knowledge = Q3.4,
         ideology = Q3.5, education = Q3.6, religiosity = Q3.7,
         gender = Q3.9, gender_other = Q3.9_4_TEXT, income = Q3.10, age = Q3.11) %>%
  # Clean up experimental condition columns
  mutate(issue = recode(issue, `human rights for refugees` = "Human rights",
                        `humanitarian assistance for refugees` = "Humanitarian assistance"),
         funding = recode(funding, `government donors` = "Government",
                          `individual, private donors` = "Private")) %>%
  mutate(crackdown = ifelse(str_detect(crackdown, "harshly restrict"), 
                            "Crackdown", "No crackdown")) %>% 
  mutate(crackdown = factor(crackdown, levels = c("No crackdown", "Crackdown"), 
                            ordered = TRUE),
         issue = factor(issue, levels = c("Human rights", "Humanitarian assistance"), 
                        ordered = TRUE),
         funding = factor(funding, levels = c("Government", "Private"), 
                          ordered = TRUE)) %>% 
  # Attention checks
  mutate(attention1_correct = str_split(Q1.3, ",") %>% 
           map_lgl(~ all(attention1_answer %in% .) & length(.) == length(attention1_answer)),
         attention2_correct = Q3.8 == attention2_answer) %>% 
  # This person contacted me separately to say that they accidentally did the first
  # attention check wrong, but that they did pay attention: 6518634
  mutate(attention1_correct = ifelse(confirmation_code == 6518634,
                                     TRUE, attention1_correct))

results <- results_check_attention %>% 
  filter(!(confirmation_code %in% codes_to_exclude$confirmation_code)) %>% 
  filter(attention2_correct) %>%
  # Factorize variables
  mutate(donate_likely = factor(donate_likely, levels = likelihood, ordered = TRUE)) %>% 
  mutate_at(vars(favor_humanitarian, favor_human_rights, favor_development),
            funs(factor(., levels = favorability, ordered = TRUE))) %>% 
  mutate(give_charity = factor(give_charity, 
                               levels = frequency_charity, ordered = TRUE),
         volunteer = factor(volunteer, levels = c("No", "Yes"), ordered = TRUE),
         political_knowledge = factor(political_knowledge, 
                                      levels = frequency_public_affairs, ordered = TRUE),
         ideology = factor(ideology, levels = levels_ideology, ordered = TRUE),
         education = factor(education, levels = levels_education, ordered = TRUE),
         religiosity = factor(religiosity, levels = frequency_religion, ordered = TRUE),
         gender = recode(gender, `Other:` = "Other"),
         gender = factor(gender, levels = levels_gender, ordered = TRUE),
         income = factor(income, levels = levels_income, ordered = TRUE),
         age = factor(age, levels = levels_age, ordered = TRUE),
         check1 = factor(attention1_correct, levels = c(TRUE, FALSE),
                         labels = levels_attention, ordered = TRUE),
         check2 = factor(attention2_correct, levels = c(TRUE, FALSE),
                         labels = levels_attention, ordered = TRUE)) %>% 
  # Dichotomize variables
  mutate(donate_likely_bin = fct_recode(donate_likely,
                                        `Not likely` = "Extremely unlikely",
                                        `Not likely` = "Somewhat unlikely",
                                        `Not likely` = "Neither likely nor unlikely",
                                        Likely = "Somewhat likely",
                                        Likely = "Extremely likely")) %>% 
  mutate_at(vars(favor_humanitarian, favor_human_rights, favor_development),
            funs(bin = fct_recode(., 
                                  `Not favorable` = "Very unfavorable",
                                  `Not favorable` = "Unfavorable",
                                  `Not favorable` = "Neutral",
                                  Favorable = "Favorable",
                                  Favorable = "Very favorable"))) %>% 
  mutate(give_charity_3 = fct_recode(give_charity,
                                     `At least once a month` = "Once a week",
                                     `At least once a month` = "Once a month",
                                     `Once a month-once a year` = "Once every three months",
                                     `Once a month-once a year` = "Once every six months",
                                     `Once a month-once a year` = "Once a year",
                                     Rarely = "Once every few years",
                                     Rarely = "Never") %>% fct_rev(),
         give_charity_2 = fct_recode(give_charity,
                                     `At least once a year` = "Once a week",
                                     `At least once a year` = "Once a month",
                                     `At least once a year` = "Once every three months",
                                     `At least once a year` = "Once every six months",
                                     `At least once a year` = "Once a year",
                                     Rarely = "Once every few years",
                                     Rarely = "Never")) %>% 
  mutate(political_knowledge_bin = fct_recode(political_knowledge,
                                              `Often` = "Most of the time",
                                              `Often` = "Some of the time",
                                              `Not often` = "Only now and then",
                                              `Not often` = "Hardly at all") %>% fct_rev()) %>% 
  mutate(ideology_3 = fct_recode(ideology,
                                 `Liberal` = "Strong liberal",
                                 `Liberal` = "Liberal",
                                 `Liberal` = "Independent, leaning liberal",
                                 `Independent` = "Independent",
                                 `Conservative` = "Independent, leaning conservative",
                                 `Conservative` = "Conservative",
                                 `Conservative` = "Very conservative") %>% fct_rev()) %>% 
  mutate(ideology_bin = fct_recode(ideology_3,
                                   `Not liberal` = "Independent",
                                   `Not liberal` = "Conservative")) %>% 
  mutate(education_bin = fct_recode(education,
                                    `No BA` = "Less than high school",
                                    `No BA` = "High school graduate",
                                    `No BA` = "Some college",
                                    `No BA` = "2 year degree",
                                    `BA and above` = "4 year degree",
                                    `BA and above` = "Graduate or professional degree",
                                    `BA and above` = "Doctorate")) %>% 
  mutate(religiosity_bin = fct_recode(religiosity,
                                      `At least once a month` = "More than once a week",
                                      `At least once a month` = "Once a week",
                                      `At least once a month` = "Once or twice a month",
                                      `Rarely` = "A few times a year",
                                      `Rarely` = "Seldom",
                                      `Rarely` = "Never",
                                      NULL = "Don't know") %>% fct_rev()) %>% 
  mutate(income_clean = fct_recode(income, NULL = "Prefer not to say"),
         income_bin = factor(income_clean >= median(.$income), levels = c(FALSE, TRUE),
                             labels = paste(c("Less than", "At least"), median(.$income)))) %>% 
  mutate(age_bin = factor(age >= median(.$age), levels = c(FALSE, TRUE),
                          labels = paste(c("Less than", "At least"), median(.$age))),
         gender_bin = fct_collapse(gender,
                                   Female = "Female",
                                   `Not Female` = c("Male", "Transgender", 
                                                    "Other", "Prefer not to say")))

# Save final clean data
saveRDS(results, here("data", "derived_data", "results_clean.rds"))
write_csvy(results, file = here("data", "derived_data", "results_clean.csv"),
           metadata = here("data", "derived_data", "results_clean.yaml"), na = "NA")
```

```{r completed-summary}
# Save information about completion rates
completed_skeleton <- results_check_attention %>% 
  left_join(codes_to_exclude, by = "confirmation_code") %>% 
  replace_na(list(reason = "Approved")) %>% 
  expand(reason, crackdown, issue, funding)

completed_summary <- results_check_attention %>% 
  left_join(codes_to_exclude, by = "confirmation_code") %>% 
  replace_na(list(reason = "Approved")) %>% 
  count(reason, crackdown, issue, funding) %>% 
  right_join(completed_skeleton, by = c("reason", "crackdown", "issue", "funding")) %>% 
  replace_na(list(n = 0))

saveRDS(completed_summary, here("data", "derived_data", "completion_summary.rds"))
write_csv(completed_summary, here("data", "derived_data", "completion_summary.csv"))
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>