### 2019-03-13 README.Rmd
Update abstract and title

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

Title <- "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
Authors <- "Suparna Chaudhry and Andrew Heiss"
Year <- "2019"
```

# `r Title`

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Department of Political Science • Christopher Newport University  
[Andrew Heiss](https://www.andrewheiss.com/) • Romney Institute of Public Service and Ethics • Brigham Young University

---

## Abstract

State crackdown on non-governmental organizations (NGOs), particularly through legal restrictions on funding and advocacy, has become increasingly pervasive across the globe. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? With reduced funding for NGOs working on contentious issues, and an absence of philanthropic culture in various developing countries, many NGOs rely on funds from private donors in Western countries. Existing research has predominantly focused on individual-level determinants of philanthropy. However, we argue that domestic political environments of NGO host countries can also influence preferences of foreign private donors, and that crackdown on NGOs serve as a heuristic to donors of organizational deservingness. Using a survey experiment, we explore how this crackdown changes donors' preferences based on the issue area and funding sources of the NGO. We find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We conclude by discussing the implications of our results on the sustainability of civil society organizations working in repressive countries.

---

This repository contains the data and code for our paper. Our pre-print is online here:

> `r Authors`. (`r Year`). *`r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at <dx.doi.org/10.17605/OSF.IO/FG53W>

## How to cite

Please cite this compendium as:

> `r Authors`. (`r format(Sys.Date(), "%Y")`). *Compendium of R code and data for `r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at <dx.doi.org/10.17605/OSF.IO/FG53W>

## How to download and install

You can either [download the compendium as a ZIP file](/archive/master.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

In order to reproduce this project, you'll need to install the compendium as an R package. After downloading the compendium, do the following:

1. Open the compendium as a project in RStudio (double click on `crackdownsphilanthropy.proj`).
2. Install the **pkgbuild**, **devtools**, **rstan**, and **rstantools** packages, either with RStudio's "Packages" panel or by running `install.packages(c("pkgbuild", "devtools", "rstan", "rstantools"))`
3. Due to [strange](https://github.com/stan-dev/rstantools/issues/52) [incompatibilties](https://github.com/klutometis/roxygen/issues/822#issuecomment-441425591) with **roxygen2** and **rstantools**, you need to compile the Stan C++ binary DLLs before building the actual **crackdownsphilanthropy** compendium package. Run `pkgbuild::compile_dll()` and wait a few minutes while the Stan models are built.
4. Finally, build the compendium package by clicking on the "Install and Restart" button in RStudio's "Build" panel.

You should now be able to run `library(crackdownsphilanthropy)` and reproduce all the findings in the paper (as well as the paper itself).

To reproduce the analysis, run `make build` from RStudio's "Terminal" panel. Open `analysis/_site/` to see the results. Run `make serve` to serve the site at <http://localhost:7000>.

To repdocuce the paper, run `make html` or `make tex` or `make docx` or `make paper` (for all three output formats) from the terminal. Open `manuscript/` (or `manuscript/tex_out/` for PDFs) to see the results.

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/))

**Code:** With only two exceptions, all the code is dual-licensed under the [MIT License](LICENSE.md) and [GPL 3](https://www.gnu.org/licenses/gpl-3.0.en.html). Two dependencies related to Stan compilation—[`R/stanmodels.R`](R/stanmodels.R) and [`tools/make_cc.R`](tools/make_cc.R)—are licensed as GPL 3 only, and we can't backwards-license them as only MIT. Additionally, any binary files generated by those files (i.e. the compiled Stan code) is inherently GPL-licensed, which also makes this entire project GPL-licensed in the end. But since we've dual-licensed pretty much everything else, you treat everything as MIT unless a file is explicitly GPL-only.

*(Here are a bunch of resources about the vagaries of mixing MIT and GPL licenses: [link](https://github.com/ropensci/unconf17/issues/32); [link](https://github.com/stan-dev/rstantools/issues/17); [link](https://opensource.stackexchange.com/questions/1640/if-im-using-a-gpl-3-library-in-my-project-can-i-license-my-project-under-mit-l))*

## Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.