### 2020-01-24 README.Rmd
Use most recent abstract

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

Title <- "Charity During Crackdown: Analyzing the Impact of State Repression of NGOs on Philanthropy"
Authors <- "Suparna Chaudhry and Andrew Heiss"
Year <- "2020"
```

# `r Title`

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Department of Political Science • Christopher Newport University  
[Andrew Heiss](https://www.andrewheiss.com/) • Andrew Young School of Policy Studies • Georgia State University

---

## Abstract

Government restrictions on NGOs have become increasingly pervasive across the globe. With reduced official aid and foundation funding available for NGOs, and an absence of philanthropic culture in many countries across the Global South, various NGOs are seeking funds from individual private donors in Western countries. How do legal restrictions on NGOs impact patterns of private philanthropic giving and individual donor decision-making? While existing research has shown that donors rely on various heuristics from social networks and organization-level attributes, we argue that domestic political environments of NGO host countries can also serve as an important structural-level heuristic that can influence donor behavior. Using a survey experiment, we find that while crackdowns do not directly influence the likelihood of donation on their own, donors are willing to donate substantially more to legally besieged NGOs abroad. We discuss the implications of our results on the sustainability of organizations working in repressive countries.

---

This repository contains the data and code for our paper. Our pre-print is online here:

> `r Authors`. (`r Year`). *`r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at <dx.doi.org/10.17605/OSF.IO/FG53W>

## How to cite

Please cite this compendium as:

> `r Authors`. (`r format(Sys.Date(), "%Y")`). *Compendium of R code and data for `r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at <dx.doi.org/10.17605/OSF.IO/FG53W>

## How to download and install

You can either [download the compendium as a ZIP file](/archive/master.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

In order to reproduce this project, you'll need to install the compendium as an R package. After downloading the compendium, do the following:

1. Open the compendium as a project in RStudio (double click on `crackdownsphilanthropy.proj`).

2. Install the **pkgbuild**, **devtools**, **rstan**, and **rstantools** packages, either with RStudio's "Packages" panel or by running `install.packages(c("pkgbuild", "devtools", "rstan", "rstantools"))`

3. You need to compile the Stan C++ binary DLLs before building and installing the actual  **crackdownsphilanthropy** compendium package ([see here for details about why](https://mc-stan.org/rstantools/articles/minimal-rstan-package.html#documentation)). Run these two lines in R:

    ```r
    pkgbuild::compile_dll()  # This will probably take a few minutes
    roxygen2::roxygenize()
    ```

4. Finally, build the compendium package by running this in R:

    ```r
    devtools::install(".")
    ```

You should now be able to run `library(crackdownsphilanthropy)` and reproduce all the findings in the paper (as well as the paper itself).

To reproduce the analysis, run `make build` from RStudio's "Terminal" panel. Open `analysis/_site/` to see the results. Run `make serve` to serve the site at <http://localhost:7000>.

To repdocuce the paper, run `make html` or `make tex` or `make docx` or `make paper` (for all three output formats) from the terminal. Open `manuscript/` (or `manuscript/tex_out/` for PDFs) to see the results.

## Data

This project includes the following data files:

- [`data/raw_data/crackdowns_philanthropy_raw.csv`](data/raw_data/crackdowns_philanthropy_raw.csv): Original results from the Qualtrics survey
- [`data/derived_data/results_clean.csv`](data/derived_data/results_clean.csv): Final version of the data, cleaned with [`analysis/01_clean-data.Rmd`](analysis/01_clean-data.Rmd)
- [`data/derived_data/results_clean.yaml`](data/derived_data/results_clean.yaml): [YAML metadata](https://csvy.org/) describing the syntax of the cleaned data
- [`data/derived_data/results_clean.rds`](data/derived_data/results_clean.rds): `.rds` version of cleaned data
- [`data/derived_data/codes_to_exclude.csv`](data/derived_data/codes_to_exclude.csv): Unique IDs of participants who failed the attention checks or took the survey outside of Mechanical Turk. These are excluded from the final data in [`analysis/01_clean-data.Rmd`](analysis/01_clean-data.Rmd)
- [`raw_data/ipums-cps/`](raw_data/ipums-cps/): Data extracts from the US Census's Current Population Survey (CPS) via the [Minnesota Population Center's Integrated Public Use Microdata Series (IPUMS)](https://cps.ipums.org/cps/):
    -  `cps_2017.dat.gz`: 2017 ASEC, with the following variables selected (in addition to whatever IPUMS preselects by default) (and weighted by `ASECWT`):
        - `AGE`
        - `SEX`
        - `EDUC`
        - `INCTOT`
    -  `cps_09_2015.dat.gz`: September 2015 basic monthly CPS (which includes the Volunteer Supplement), with the following variables selected (and weighted by `VLSUPPWT`):
        - `VLSTATUS`
        - `VLDONATE`

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/))

**Code:** With only two exceptions, all the code is dual-licensed under the [MIT License](LICENSE.md) and [GPL 3](https://www.gnu.org/licenses/gpl-3.0.en.html). Two dependencies related to Stan compilation—[`R/stanmodels.R`](R/stanmodels.R) and [`tools/make_cc.R`](tools/make_cc.R)—are licensed as GPL 3 only, and we can't backwards-license them as only MIT. Additionally, any binary files generated by those files (i.e. the compiled Stan code) is inherently GPL-licensed, which also makes this entire project GPL-licensed in the end. But since we've dual-licensed pretty much everything else, you treat everything as MIT unless a file is explicitly GPL-only.

*(Here are a bunch of resources about the vagaries of mixing MIT and GPL licenses: [link](https://github.com/ropensci/unconf17/issues/32); [link](https://github.com/stan-dev/rstantools/issues/17); [link](https://opensource.stackexchange.com/questions/1640/if-im-using-a-gpl-3-library-in-my-project-can-i-license-my-project-under-mit-l))*

## Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.