### 2020-06-17 manuscript/manuscript.Rmd
My R&R changes

---
title: "Dynamics of international giving: Exploring individual donor preferences"
short-title: Charity During Crackdown
author:
- name: Suparna Chaudhry
  affiliation: Lewis \& Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "`r format(Sys.time(), '%B %e, %Y')`"
git-repo: https://github.com/andrewheiss/ngo-crackdowns-philanthropy
reference-section-title: References
toc: false
title-page: true
abstract: >-
  Government restrictions on NGOs have become increasingly pervasive across the globe. While this crackdown has been shown to have a negative impact on public funding flows, we know little about how it impacts private funding flows. How does information about crackdown abroad, as well as organizational attributes of nonprofits affect individual donors’ willingness to donate internationally? Using a survey experiment, we find that learning about repressive NGO regulatory environments increases generosity in that already-likely donors are willing to donate substantially more to legally besieged nonprofits. This generosity persists when mediated by two organizational-level heuristics: NGO issue areas and its main funding sources. We conclude by discussing the implications of our results on the scholarship and practice of nonprofits using different framing appeals to increase fundraising at a time when traditional funding to such organizations is decreasing. 
thanks: >-
  We thank Mark Buntaine, Michael DeCrescenzo, Jeffrey Friedman, Thomas Leeper, and Tristan Mahr for their helpful comments and input. We also thank Samantha Camilletti and Joshua Dutro for their excellent research assistance. This experiment received approval by the human subjects research committees at both Christopher Newport University (018.042) and Brigham Young University (E18104). We preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/dx973/>.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Lewis \& Clark College; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
additional-info: Figures should appear in color online only, not in print.
correspondence: >-
  Correspondence concerning this article should be addressed to Suparna Chaudhry, PhD, Department of International Affairs, Lewis \& Clark College, 0615 SW Palatine Hill Rd. Portland, OR 97219. Contact: schaudhry@lclark.edu
word-count: "FILL"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression
---

<!-- manuscript.md is generated from manuscript.Rmd. Only edit the .Rmd file, *not* the .md file. -->

```{r setup, include=FALSE}
library(tidyverse)
library(pander)
library(scales)
library(glue)
library(here)

source(here("analysis", "options.R"))

# Formatting functions
tex_percent <- percent_format(accuracy = 0.1, suffix = "\\%")  # Escape for TeX
percent_0 <- percent_format(accuracy = 1)  # 5%
percent_1 <- percent_format(accuracy = 0.1)  # 5.5%
md_dollar <- dollar_format(prefix = "\\$")  # Escape for Markdown/TeX

# Load and clean data
likely <- readRDS(here("data", "derived_data", "results_models_likely.rds")) %>%
  # Make LaTeXy versions of numbers
  mutate(delta = glue("$\\Delta = {x}$", x = tex_percent(theta_diff)),
         p_neq0 = glue("$p(\\Delta \\neq 0) = {x}$", x = round(p.diff.not0, 2))) %>%
  # Simplify title
  mutate(title = str_replace_all(title, " \\| Crackdown", "")) %>%
  # Add subscripts to funding and issue numbers
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Funding",
                         str_replace(., "Delta", "Delta_\\\\text{funding}"), .))) %>%
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Issue",
                         str_replace(., "Delta", "Delta_\\\\text{issue}"), .)))

amount <- readRDS(here("data", "derived_data", "results_models_amount.rds")) %>%
  # Make LaTeXy versions of numbers
  mutate(delta = glue("$\\Delta = {x}$", x = round(mu_diff, 2)),
         p_neq0 = glue("$p(\\Delta \\neq 0) = {x}$", x = round(p.diff.not0, 2))) %>%
  # Simplify title
  mutate(title = str_replace_all(title, " \\| Crackdown", "")) %>%
  # Add subscripts to funding and issue numbers
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Funding",
                         str_replace(., "Delta", "Delta_\\\\text{funding}"), .))) %>%
  mutate_at(vars(delta, p_neq0),
            list(~ifelse(title == "Issue",
                         str_replace(., "Delta", "Delta_\\\\text{issue}"), .)))

results <- readRDS(here("data", "derived_data", "results_clean.rds"))
```

Over the last three decades, governments across the world have sought to limit the work of nonprofits and nongovernmental organizations (NGOs)—a phenomenon commonly known as "closing civic space" [@Carothers:2015; @DPR:2015]. Repressive governments have enacted laws creating barriers to advocacy, funding, and entry, adversely affecting the work of thousands of NGOs, which typically rely on funding from donors across the world. This is particularly the case in countries with weak institutions and poor governance, where international donors typically seek out NGOs to implement their projects, as direct aid transfers to recipient governments pose the risk of misuse and bureaucratic inefficiency [@Dietrich2013]. As a result of this global crackdown on civil society, however, official aid flows to NGOs have decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. While we know there has been a negative impact on funding from official governmental donors, we know relatively little about how this crackdown affects private donors, in particular how it influences individual donors' willingness to engage in philanthropy.

Charitable giving to nonprofits working in international affairs has steadily increased over the past two decades, with a noticeable increase over the last decade. Between 2014 and 2016, giving to international affairs-focused issues increased by 14.1% [@givingusa2017report]. This amount is not insignificant—by 2016, total private giving to civil society organizations working in international affairs had increased to \$22.03 billion [@givingusa:2017]. Most notably, a large portion of this growth is driven by individuals making small-scale donations. In 2017, individuals gave $286.7 billion, or approximately 70% of total giving to international affairs. This further increased to $427.71 billion in 2018, 85.7% of which was through individuals [givingusa:2019]. Much of the existing research on donations to international causes is restricted to elite, high net-worth donors [@USTrust2014], but these large amounts demonstrate that it is imperative to understand the motivations and attributes of more ordinary individual donors. What factors affect individual donor preferences on giving to international causes? More specifically, how does information about regulatory environments abroad, as well as organizational-level attributes of NGOS— such as their main issues of focus and sources of funding—affect donors' willingness to donate internationally?

Given information asymmetry and time constraints, individual donors often use heuristics to simply their decision-making [@TverskyKahneman:1974; @CrosonShang:2011; @Tremblay-BoirePrakash:2017], We leverage an online survey experiment using Amazon's Mechanical Turk to assess how different structural and organizational-level heuristics affect individual donors' willingness to donate. We find that learning about more repressive NGO regulatory environments increases generosity in that already likely donors are willing to donate substantially more to legally restricted NGOs.

This generosity persists when mediated by two important organization-level heuristics: NGO issue area and funding sources. Learning about crackdown increases individuals' willingness to donate to privately funded nonprofits. This may be because the survival of the NGO appears to be in question and the donor's assessment might be that their donation actually makes a difference. These results are particularly substantive for privately-funded human rights NGOs facing crackdown, with donors showing an increased willingness to not just donate to them, but also to donate more to them. While not unsurprising, this does suggest that private donors to human rights NGOs likely know that their work is always challenging to host governments, and that such organizations need greater support when facing stricter regulatory environments.

Our study makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individuals' willingness to donate. First, while factors shaping individual giving to domestic causes have been thoroughly explored in existing literature, the dynamics of international giving and individual donor preferences are less explored.[^exceptions] Understanding the factors driving individual giving is particularly important when when looking at international issues, as philanthropy to international causes can be harder to motivate because of the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100].

Second, most nonprofits working internationally have traditionally relied on government and foundation funding. But in the era of closing civic space, these groups may need to tailor their fundraising strategies to individual donors. Framing, or the process through which actors present information to influence perceptions of behavior, is particularly important in this regard.[^framing] However, we lack systematic studies of which frames may be effective when making appeals to donors considering international philanthropy. Our study examines the impact of political frames such as regulatory crackdown, and how these frames interact with organizational attributes to signal NGO deservingness. Accordingly, our results can help international nonprofits frame their appeals for funding, especially when facing restricted legal space abroad.

Below, we summarize existing research on the determinants of giving to NGOs and lay out our expectations regarding how crackdowns on NGOs may affect how individuals engage in philanthropy. We then describe our survey experiment and present the results. We conclude with implications for NGOs working in repressive countries and lay out questions ripe for future research.

# Philanthropy towards international NGOs

Thousands of international NGOs in the Global South receive funds from a variety of public and private sources each year.[^NGO] Though our focus in this paper is one private funding, it is essential to differentiate between the two sources. Public aid (more commonly known as traditional donor aid) channels funds from official aid agencies towards nonprofits through a variety of bilateral and multilateral institutions. Private funds—which are comparatively newer international donors compared to traditional donor aid—can come from foundations, private corporations, and individuals.

In this paper, we specifically look at individual-level determinants to international philanthropy. Private giving to nonprofits working in international affairs has increased consistently over the past decade.[^internationalaffairs] Moreover, donations from individuals have far outstripped foundation and corporate giving—in 2018, individual donations amounted to \$292.09 billion, compared to \$75.86 billion from foundations and \$20.05 billion by corporations. While foundations are often seen as the primary funders of international charities, donations from foundations accounted for only 18% of all charitable giving in 2018—and grants from the Bill and Melinda Gates Foundation accounted for more than half of all foundation giving [@givingusa:2019]. Individual donors play an incredibly important—and underexplored—role in funding international NGOs.

Given these substantive amounts, understanding individual-level determinants of giving to international nonprofits is important to investigate as individual private donors may not necessarily mirror the preferences of official aid agencies. While there is evidence that citizens generally support the goals of aid agencies [@MilnerTingley2010], surveys show that the American public is notoriously misinformed about US foreign aid [@norris:2017]. Further, both foreign aid and democracy aid are often considered unresponsive to political developments in recipient countries [@brainard2007; @Carothers:2015] and public aid decisions—especially about final recipients and amounts—are not easily accessible. There may also be heterogeneity within how donor governments allocate funds. For instance, unlike established donors like the United States and other OECD countries, emerging donor governments tend to demand fewer human rights, governance, or environmental preconditions [@fenglerkharas].

More recent research on the allocation of private aid flows has also found little support for the idea that individual donor behavior mimics that of official donors. @desaikharas find that unlike official donors who tend to channel funds to countries with better institutional quality or as a reward for governmental performance [@Bermeo:2011; @DietrichWright:2015], individual donors do not use the same performance-based metrics. Rather, individuals are more inclined to donate to groups in countries experiencing a humanitarian crisis—in particular, crises following natural disasters. @desaikharas conclude that compared to donor agencies, "private donors respond to different project and country characteristics" (p. 517) Given these above reasons, the average individual donor will not follow the same decision-making logic as donor governments, and individual international philanthropy should therefore be investigated on its own terms.

Understanding the motivations of individual private donors is also essential because private funding flows towards nonprofits working internationally appear to not be as adversely impacted as public funding flows in the face of the global crackdown on NGOs. Prior research has demonstrated that restrictions on cross-border funding has reduced official aid and official development assistance to repressive countries [@citeme]. But data from the Foundation Center shows no correlation between amount of funding flows from US private foundations to a given country and that country's score on the Index of Philanthropic Freedom. Additionally, preliminary evidence shows that "US foundations remain committed to channeling funds to countries where the legal environment is growing increasingly unfriendly." [@citeme-too]. As such, it is imperative to understand whether individual donors mirror the trajectory of foundations when giving internationally, especially to countries with restrictive legal environments.

## Key drivers of individual-level philanthropy

Substantial research on charitable giving has looked at the motivations of private donors, with much of this work focusing on the individual-level determinants of giving [@Wiepking:2010; @BekkersWiepking:2011]. However, research on individual donor behavior has overwhelming examined giving to local—rather than international—causes. This literature has primarily focused on the importance of three main factors, especially when considering charitable behavior towards international causes: (1) the role of social capital and social networks; (2) the role of individual experiences such as higher education and levels of religiosity; and (3) third party certifications.

One major area of inquiry in philanthropic studies looks at the role of social capital, or the "networks, norms, and social trust that facilitate coordination and cooperation for mutual benefit" [@Putnam:1995, 67]. While quite broad, the literature concentrates on "the networks of formal and informal relationships to which people are associated," [@havens2007, 240] or associational capital. Prior research has found that those who participate in a variety of associations are more likely to make larger donations [@hossainlamb2017]. Further, those who volunteer their time (compared to those who did not), were both more likely to make charitable donations and make bigger donations, since volunteering exposed individuals to the need for charitable donations and made philanthropy more compelling [@hossainlamb2012]. An individual's social networks also play an important in deciding to donate. Looking at relational configurations in social networks, @HerzogYang:2018 find that having both a giver and a solicitor in one's social network increase the likelihood of that individual's charitable giving.

Individual experiences such as access to higher education and increased religiosity can also influence philanthropic giving. Higher education fosters prosocial motivations and brings people into particular social networks that entail a higher level of solicitations [@BekkersWiepking:2011]. Education also increases abstract thinking [@wiepkingmaas, 1978], which is particularly important for donating to more distant international relief organizations than donating to local, more visible nonprofits [@Micklewright:2009]. Levels of religiosity are also a significant predictor of the likelihood and level of giving. People who regularly attend religious services are more likely to have prosocial values, face a stronger norm of charitable giving, and may feel pressure to conform to group and community standards [@wiepking2010]. Overall, individuals with higher levels of income, education, and greater religious proclivities have been shown to not only be more likely to give internationally, but also to give higher amounts [@Rajan:2009; @CasaleBaumann:2015].

Finally, information about organizational characteristics and third party certifications can also impact levels of individual philanthropy, through results are mixed. Existing research shows that information on the efficiency of NGOs merely steers rather than encourages or discourages overall donations [@ryazanov:efficiency]. The effects of changes in charity watchdog ratings and nonprofit accountability systems on individual donors is also mixed, showing that donors do not have the time to research charities thoroughly [@Bekkers:2010] and that ratings by charity watchdogs do not really affect donor support for these nonprofits [@SzperPrakash:2011].

While this robust literature looks at the drivers of individual-level philanthropy, it overwhelmingly focusses on giving to local rather than international causes. This domestic focus is warranted, though: on average, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. However, this focus on domestic philanthropy causes means that that the precise causal mechanisms through which cross-border charity can be motivated and mobilized is still under-examined [@desaikharas, 517]. This problem is further compounded by the fact that there is greater paucity of data and diversity of sources regarding international philanthropy. Indeed, most research on individual-giving to international causes has been restricted to elite, high net-worth donors [@USTrust2014].[^highnetworth] Given the increasing importance of more ordinary individual donations for international NGOs, it is crucial to better understand what motivates private donors to engage in internationally-focused philanthropy.

As explored previously, individual private donors do not necessarily mirror the choices made by public donors. Individuals rely on signals and heuristics that differ from the benchmarks used by large governmental agencies and private foundations. These signals and heuristics, we argue, can be even more important in giving to international causes. In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking complete information about an organization's deservingness and efficiency is costly and time-consuming. These heuristics are commonly rooted in donor perceptions of organizational characteristics—donors make cursory judgments about an organization's issue area, mission, vision, and values, and seek out supplementary information from friends, family, and acquaintances [@SzperPrakash:2011; @Sloan:2009]. Nonprofits fundraising internationally may be therefore more reliant on framing. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

# Theorizing the effects of political and organizational frames on individual giving

In addition to organizational factors that signal an NGO's trustworthiness and deservingness, we argue that structural factors—including whether an organization works abroad and the domestic political environment of the group's host country—can also serve as important heuristics in the decision to engage in philanthropy [@CasaleBaumann:2015; @KnowlesSullivan:2017; @Tremblay-BoirePrakash:2017]. We propose that the regulatory relationship between international NGOs (INGOs) and their host governments serves as a signal of NGO trustworthiness and deservingness. NGO legislation is not inherently restrictive, and governments frequently use laws to regulate the behavior of NGOs [@BloodgoodTB; @DeMattee:2018]. Restrictive NGO laws, on the other hand, are designed to limit organizational programming by imposing barriers to entry, funding, and advocacy. These stricter anti-NGO laws can signal to donors that governments perceive these groups as threatening and may crack down on them.

We argue that donors perceive legal crackdowns differently from organization-level heuristics—such as NGO missions, visions, values—because donors may be able to influence these latter attributes, at least to some extent. Conversely, donors can do little to change the domestic political environment of another country. Legal crackdowns abroad can signal to donors that the government may eventually target other NGOs as well, which can increase donor urgency to support legally besieged NGOs, even if donors are not completely satisfied with some organization-level features of a group. As such, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization and increase the amount donated.

> **H~1a~**: If donors learn that NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.
>
> **H~1b~**: If donors learn that NGOs face legal crackdowns abroad, then individual donors will donate more to them.

@fig:causal-path represents this relationship formally. Legal crackdowns (**C**) have a direct effect on donor decisions (**Y**; donation likelihood and amount) because of their heuristic function. Legal crackdowns do not occur in isolation, however. Countries are more likely to repress NGOs both when organizational issue areas (**I**) are in tension with government preferences and when organizations receive substantial funding (**F**) from abroad [@DPR:2015]. Both of these organizational characteristics thus confound the causal effect of crackdowns on donor decisions and must be accounted for in the analysis—issue areas and funding sources simultaneously influence the likelihood of crackdown while also serving as organizational heuristics that influence donor perceptions of NGO deservingness.

![Causal diagram of the relationship of organizational and structural causes of donor behavior](output/figures/causal-path.pdf){#fig:causal-path}

The *contentiousness of NGO issue areas*—or the degree to which an NGO's programming is compatible with government preferences, can also serve as an organization-level heuristic.[^targeting] NGOs that address humanitarian issues such as relief and development have broader appeal to donors and these issues rarely challenge the government. Further, their programmatic output is easily quantifiable and donors can see the results of their charity more readily [@Bush:2015]. Previous research has shown that that public donors do indeed differentiate between NGO issue areas, especially when these organizations are working in repressive contexts. In particular, donor agencies in the Organisation for Economic Cooperation and Development (OECD) respond to repressive NGO environments, especially those containing barriers to advocacy, by decreasing funds for groups working on politically sensitive causes such as anti-corruption initiatives, elections, human rights, legal reform and security sector reform, and instead increasing funding for groups working on relatively tamer causes such as health, education and agriculture [@ChaudhryHeiss:2018].

Issue area contentiousness directly influences legal crackdowns, however, and confounds the heuristic effect of crackdowns on donor behavior (see @fig:causal-path). NGOs working on more contentious human rights issues can be challenging to regimes [@Heiss:2017]. These organizations are often viewed as politically motivated outsiders and run a higher risk of getting expelled [@DPR:2015], increasing the possibility of wasting donor resources. There is evidence that this a concern for private donors. In a survey of women’s NGOs across the globe conducted by Mama Cash and Urgent Action Fund, many NGOs argued that private donors (especially foundations) were withdrawing from funding sensitive issues in repressive contexts, which in turn hurt women and trans rights organizations. Some private funders, these groups argued, were also more likely to fund well-established groups that were more likely to survive crackdown.

However, compared to our existing knowledge on public donors and foundation donors, we lack knowledge about how individual donors respond to frames emphasizing one nonprofit issue area over another, especially a contentious area over a tamer one. NGOs working on tamer issues such as health, education, disaster relief, and humanitarian aid can often frame their activities as in need of greater support—the perception of need, whether through personal experiences or media framing. This may strongly influence individual decisions to donate to relief efforts [@wiepking2010]. Individual donors may believe that it is more difficult to improve human rights in the host country and subsequently shy away from donating to NGOs working on contentious causes.

As such, we hypothesize that donors will be more likely to donate to less contentious NGOs because these are seen as uncontroversial, apolitical, directed towards the most deserving of need, and are more capable of producing short-term quantifiable results. We use humanitarian NGOs as a proxy for NGOs working on non-contentious issues and human rights NGOs as a proxy for contentious issues.

> **H~2a~**: If donors learn that legally besieged NGOs work on humanitarian issues, then individual donors will be more likely to donate to them.
>
> **H~2b~**: If donors learn that legally besieged NGOs work on humanitarian issues, then individual donors will donate more to them.

The final heuristic we examine is the *source of NGO funding*. Government grants to nonprofits can often crowd out private funding, though evidence is mixed on whether this completely or partially crowds out private dollars [@Steinberg:1991; @Heutel:2014]. Donors may feel less inclined to contribute to NGOs receiving funds from agencies like USAID because such funding can signal a lack of independence from the government. Organizations that receive substantial funding from their home governments frequently avoid programming that would question the donor government even if such programming is in line with their mission [@Stroup:2012]. In contrast, individual donors who see that organizations are privately funded may feel that they can also contribute and help and that their marginal donation would make a noticeable difference.

As funding restrictions are one of the most common forms of anti-NGO regulations [@DPR:2015], organizational funding again confounds the effect of anti-NGO crackdowns on donor behavior (see @fig:causal-path). Accordingly, the combination of legal crackdown and funding may reshape donor perceptions of deservingness. Individual donors may be less likely to donate to government-funded NGOs that face restrictions, since legal trouble with the organization's primary funder could fuel private donor mistrust (i.e. perhaps the NGO did something to deserve crackdown). Donors thus may be likely to donate more to NGOs that rely on private funding.

> **H~3a~**: If donors learn that legally besieged NGOs receive substantial funding from government sources, then individual donors will be less likely to donate to them.
>
> **H~3b~**: If donors learn that legally besieged NGOs receive substantial funding from government sources, then individual donors will donate less to them.

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Donors do not always know about anti-NGO laws in host countries or who funds an NGO when they make their contributions. Moreover, the average donor will not know exactly what "legal crackdown" entails for any given organization (i.e. does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguous definition is not a central concern as the heuristics we explore here are explicitly simplified shortcuts to donor information-gathering. Knowing that an organization works with human rights issues (even if donors are not familiar with specific kinds of rights) should be enough to trigger the issue-based deservingess heuristic. Similarly, knowing that an NGO faces legal trouble should be enough to trigger the structural deservingness heuristic. Moreover, while many NGOs and nonprofits make information about their funding sources easily accessible through annual reports, donors often do not take the time to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, issue area, and funding sources. Instead, donors rely on signals in basic information about the organization to shape their preferences and behavior.

We also posit that these signals work interactively to shape donor behavior and that combinations of characteristics will have different effects. For instance, the effect of crackdown will be different depending if an organization focuses on humanitarian assistance or human rights, or if it receives government funding or not. Because it is unclear a priori what these interactive effects will be, we test nested versions of the hypotheses in our analysis.

# Research design

To measure the effect of structural and organization-level heuristics on individual donor preferences, we use an experiment to vary the different signals donors are exposed to. Comparing these effects allows us to measure the relative strength of these heuristics, and ultimately help NGOs tailor their fundraising strategies for private individuals. Prior to launching the experiment, we preregistered our hypotheses and research design at the Open Science Framework, and our preregistration protocol is available at [https://osf.io/dx973/](https://osf.io/dx973/). We made two minor deviations from the original preregistration, described in detail in the online appendix.[^deviations] The online appendix also includes the full text of the survey experiment, as well as details about our sample demographics, balance across experimental conditions, and Bayesian priors. The analysis is fully reproducible using code and data available at ANONYMIZED.

## Sample

We test these hypotheses with a vignette-based survey experiment fielded through Amazon's Mechanical Turk.[^mturk] Our target population is the portion of Americans hypothetically willing to donate money for human rights and humanitarian work abroad. Our convenience sample (March 2018, N = `r nrow(results)`) generally approximates the characteristics of our target population, since it is younger, more educated, wealthier, and more likely to donate to charities than nationally representative samples. Additionally, the majority of the sample (≈90%) feels favorably toward human rights, humanitarian, and development NGOs.

## Experimental treatments and outcomes

We presented participants with a short paragraph with three manipulated frames, each highlighting different factors that might influence patterns of private philanthropy: (1) crackdown/no crackdown, (2) humanitarian assistance/human rights, and (3) government/private funding. We used the International Rescue Committee (IRC) as our example NGO because its work easily applies to each possible frame. IRC provides humanitarian assistance and engages in human rights advocacy for refugees, more than a quarter of IRC's income comes from both government grants and private donations, and it works in countries with and without anti-NGO laws. We use a 2×2×2 between-subject factorial design with participants randomly assigned to one of eight versions of the following vignette:[^consort]

> The International Rescue Committee (IRC) focuses on *{humanitarian assistance for refugees | human rights for refugees}* *{and works in countries that have recently passed laws that harshly restrict nonprofit organizations | NOTHING}*. A substantial proportion of IRC's funding comes from *{government | private}* donors.

We measured two outcomes: (1) how likely participants would be to donate to IRC (measured with a 5-point scale ranging from "Extremely likely" to "Extremely unlikely"), and (2) how much participants would hypothetically donate to IRC if they had an extra $100. We collapsed the likelihood scale into a binary variable measuring whether the participant is likely ("Extremely likely" and "Somewhat likely") or not likely (all other responses) to donate.[^collapse] We also included an exploratory free response question asking respondents to justify their choice.

Each of the varied treatments correspond to our hypotheses: crackdown vs. no crackdown, humanitarian assistance vs. human rights, and government vs. nongovernmental funding. Participants might misinterpret the vignette and assume that IRC's host governments fund the organization rather than its home government. In either case, knowing about government funding should still serve as a signal of the organization's deservingness. Similarly, there could be confusion with phrase "private funding", but since we use it as the opposite of government funding, it should still act as a heuristic. Free responses tend to confirm the correct interpretation of the vignette, and many respondents explicitly justified their lack of support for IRC because of government funding, as expected ("I see that a large portion of its funding comes from government donors, so I feel it doesn't need me as much."[^6216778])

## Estimation

We test our first hypothesis by calculating the differences in the average likelihood to donate and the average amount donated across the crackdown conditions. For our second and third hypotheses we measure the difference in means spread across both issue and funding conditions. For additional exploration of the effect of issue and funding, we also measure the effect of crackdown within nested combinations of issue and funding conditions.

We use two Bayesian models to estimate the effect of crackdown on the likelihood of donating and the amount hypothetically donated. We model the proportion of respondents indicating they would likely donate as a binomial distribution:

\footnotesize

$$
\begin{aligned}
n_{\text{group 1, group 2}} &\sim \text{Binomial}(n_{\text{group total}}, \theta_{\text{group}}) &\text{[likelihood]}\\
\Delta_n &= n_{\text{group 2}} - n_{\text{group 1}} &\text{[difference in proportions]} \\
n &: \text{Number likely to donate} \\
\\
\theta_{\text{group 1, group 2}} &\sim \text{Beta}(5, 5) &\text{[prior prob. of likelihood]}
\end{aligned}
$$

\normalsize

We estimate the mean amount donated in each condition using a *t* distribution:

\footnotesize

$$
\begin{aligned}
x_{\text{group 1, group 2}} &\sim \text{Student } t(\nu, \mu, \sigma) &\text{[likelihood]}\\
\Delta_x &= x_{\text{group 2}} - x_{\text{group 1}} &\text{[difference in means]} \\
x &: \text{Mean amount donated} \\
\\
\nu &\sim \text{Exponential}(1 / 29) &\text{[prior normality]} \\
\mu_{\text{group 1, group 2}} &\sim \mathcal{N}(\bar{x}_{\text{group 1, group 2}}, 10) &\text{[prior donation mean per group]}\\
\sigma_{\text{group 1, group 2}} &\sim \text{Cauchy}(0, 1)&\text{[prior donation sd per group]}
\end{aligned}
$$

\normalsize

We use median values from the posterior distributions as point estimates and calculate credible intervals using the 95% highest posterior density. We declare an effect statistically significant if the posterior probability of being different from zero exceeds 0.95. Because differences can be either positive or negative (i.e. in some frames, the crackdown condition causes larger donations; in others it causes smaller donations), we report the probability that the difference is not equal to zero: when the median value is negative we report the proportion of predicted values that are negative, and vice versa.

# Results

@tbl:hypotheses provides a summary of how our results map onto our hypotheses, both individually and nested within each other.

\stgroup

\renewcommand*{\arraystretch}{1.35}

```{r hypothesis-summary, warning=FALSE, message=FALSE, echo=FALSE, results="asis"}
read_csv(here("data", "manual_data", "hypothesis-summary.csv")) %>%
  mutate_at(vars(Hypothesis, Expectation),
            ~str_replace_all(., "zzz", "\\\\\\\n")) %>%
  pandoc.table(keep.line.breaks = TRUE, style = "grid", justify = "llll",
                      caption = "Summary of hypotheses and results {#tbl:hypotheses}")
```

\fingroup

## Likelihood of donation

Without considering issue or funding, we find that respondents are not more likely to donate to legally besieged INGOs (H~1a~). As seen in @fig:likely-diffs(A), those exposed to the crackdown condition tend to have a slightly higher probability of donating (`r percent_1(filter(likely, title == "Crackdown")$theta_2)`) than those in the control condition (`r percent_1(filter(likely, title == "Crackdown")$theta_1)`), but the difference is not significant (`r filter(likely, title == "Crackdown")$delta`; `r filter(likely, title == "Crackdown")$p_neq0`). We find similar results when looking at the isolated effects of issue area and funding. Both differences are positive, but the posterior probability of being larger than zero is low (`r filter(likely, title == "Issue")$delta`; `r filter(likely, title == "Issue")$p_neq0`; `r filter(likely, title == "Funding")$delta`; `r filter(likely, title == "Funding")$p_neq0`). We thus do not initially find evidence that crackdowns (or issue or funding) alone increase the likelihood of donating.

![Difference in likelihood of donation across crackdown and no crackdown groups, conditioned by other experimental frames](output/figures/likely-diffs.pdf){#fig:likely-diffs}

```{r child = "output/tables/tbl-likely-diffs.md"}
```

@fig:likely-diffs(B–C) shows the difference in the likelihood of donation for both crackdown conditions across NGO issue areas (H~2a~) and funding sources (H~3a~). Crackdown has almost no effect on the likelihood of donating to human rights NGOs (`r filter(likely, title == "Human rights")$p_neq0`), which is not surprising, since people who donate to these kinds of NGOs likely know that their work is challenging to host governments. We hypothesized that the positive crackdown effect would be stronger for humanitarian NGOs, and while crackdown does have a positive effect under the humanitarian frame, the probability this difference is greater than zero is below our 95% threshold (`r filter(likely, title == "Humanitarian assistance")$p_neq0`). We do find some evidence for H~3a~, however. Crackdown has little effect on the likelihood of individual preferences to donate to NGOs receiving government funding (`r filter(likely, title == "Government")$p_neq0`), but it does increase the likelihood of donating to privately funded NGOs by `r percent_0(filter(likely, title == "Private")$pct_change)` (`r filter(likely, title == "Private")$delta`; `r filter(likely, title == "Private")$p_neq0`), possibly because a donor might feel that their marginal donation would make a difference.

While the crackdown frame has no effect on the probability of donating to human rights NGOs alone, conditioning this finding on the source of NGO funding reveals competing trends. @fig:likely-diffs(D) shows the difference in donation likelihood across all experimental conditions. When donors know that a human rights organization is primarily government-funded, they are substantially less likely to donate if the organization faces government crackdown abroad, which follows our expected hypotheses for both issue area and funding (`r filter(likely, title == "Human rights | Government")$delta`; `r filter(likely, title == "Human rights | Government")$p_neq0`), and could possibly reflect donor suspicion about the mission and goals of government-funded human rights programming. In contrast, when donors know that a human rights organization is privately funded, they are more likely to donate when there is a legal crackdown than when the NGO faces no legal hurdles (`r filter(likely, title == "Human rights | Private")$delta`; `r filter(likely, title == "Human rights | Private")$p_neq0`). The private funding heuristic appears to override the dampening in likelihood that should come from focusing on more contentious human rights issues. For humanitarian assistance NGOs, however, legal crackdowns do not have a significant effect on the likelihood of donation when nesting either funding and issue frames—neither of the nested experimental frames play a heuristic role.

These results give partial support to H~3a~. The source of funding has little influence on the preference to donate to humanitarian NGOs, but on average, donors are substantially more likely to want to give to privately funded human rights NGOs than government-funded organizations. Additionally donors appear to both punish government-funded human rights NGOs facing crackdown and rally behind privately funded NGOs facing those same hurdles. For instance, one respondent who was presented with a government-funded human rights version of IRC explained that "I'm not exactly sure what they did was right or wrong, I think it just seems bad, so I'm less likely to want to donate to them."[^8425616] Other respondents showed increased support when presented with a privately-funded version of IRC, explaining that they would donate a substantial amount because "[t]hey [IRC] are doing good work in countries where it is tough for groups like them to operate and they need all the help they can get."[^3025921] Individual donors thus seem to be more willing to support besieged human rights organizations when they are unencumbered by government funds.

## Amount donated

On their own, crackdowns do not substantially influence donors' *likelihood* to donate, but they do increase the *amount* of money that respondents are willing to contribute (see @fig:amount-diffs(A)). Informing participants that IRC faces legal hurdles abroad increased donations by `r md_dollar(filter(amount, title == "Crackdown")$mu_diff)` on average, a `r percent_0(filter(amount, title == "Crackdown")$pct_change)` increase from the no crackdown condition (`r filter(amount, title == "Crackdown")$p_neq0`). This provides good evidence for H~1b~—donors appear to give more money to besieged INGOs. In contrast, neither issue area nor funding on their own have an effect on the amount donated (`r filter(amount, title == "Issue")$delta`; `r filter(amount, title == "Issue")$p_neq0`; `r filter(amount, title == "Funding")$delta`; `r filter(amount, title == "Funding")$p_neq0`).

![Difference in amount donated across crackdown and no crackdown groups, conditioned by other experimental frames](output/figures/amount-diffs.pdf){#fig:amount-diffs}

```{r child = "output/tables/tbl-amount-diffs.md"}
```

This trend also holds when nesting crackdown within NGO issue area. The crackdown condition elicits higher donations for both the human rights and humanitarian NGOs, though with varying levels of significance (see @fig:amount-diffs(B–C)). Emphasizing legal crackdown increases donations to human rights NGOs by `r md_dollar(filter(amount, title == "Human rights")$mu_diff)`, but with a lower probability of significance (`r filter(amount, title == "Human rights")$p_neq0`). In contrast, crackdowns boost donations to humanitarian NGOs by `r md_dollar(filter(amount, title == "Humanitarian assistance")$mu_diff)`, a `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` increase beyond the control condition (`r filter(amount, title == "Humanitarian assistance")$p_neq0`). We thus find support for H~2b~: donors give more to besieged humanitarian NGOs. We also find evidence for H~3b~. Donors give `r md_dollar(filter(amount, title == "Private")$mu_diff)` more (a `r percent_0(filter(amount, title == "Private")$pct_change)` increase) to legally besieged NGOs that are privately funded, as hypothesized, (`r filter(amount, title == "Private")$p_neq0`), while the crackdown condition has little effect on the amount donated to government-funded NGOs (`r filter(amount, title == "Government")$p_neq0`), possibly because the effects cancel each other out—legal crackdowns should increase the amount donated, but reliance on government funding dissuades donors from giving.

Combining issue areas and funding sources provides more texture. We previously found that donors less likely to donate to government-funded human rights NGOs and more likely to donate to privately funded human rights NGOs when they face legal issues. These trends also apply to preferences in the amount donors donate. Respondents were willing to donate `r md_dollar(filter(amount, title == "Human rights | Private")$mu_diff)` more to besieged privately funded human rights NGOs compared to the no crackdown control condition (`r filter(amount, title == "Human rights | Private")$p_neq0`), increasing their donations by `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)`. Respondents explained that they chose to give additional money to privately funded human rights NGOs precisely because "the country limits non-profits."[^13685] The punishment mechanism found previously does not translate to the amount donated—while there is a decrease of `r md_dollar(abs(filter(amount, title == "Human rights | Government")$mu_diff))` in average donations to government-funded human rights NGOs facing crackdown, the chance this difference is less than zero is lower than our threshold (`r filter(amount, title == "Human rights | Government")$p_neq0`).

Emphasizing legal difficulties increased donations to government-funded humanitarian NGOs by `r percent_0(filter(amount, title == "Humanitarian assistance | Government")$pct_change)`, or `r md_dollar(abs(filter(amount, title == "Humanitarian assistance | Government")$mu_diff))` (`r filter(amount, title == "Humanitarian assistance | Government")$p_neq0`). The crackdown condition causes this increase despite some hesitation from respondents who often indicated that government-funded NGOs do not deserve private donations. Respondents reasoned since IRC "is already receiving funding from governments, so funding from private individuals doesn't seem as necessary. I'd prefer to give my money to an organization that is primarily run by private donations"[^6486084] or because "[m]y tax paying dollars go to the government and if the government is helping to fund it then they are in turn getting some of my money."[^9509936] This hesitation is not universal, however, and in aggregate most respondents chose to give substantially more. Many echoed this respondent, who identified the crackdown condition as one of the reasons for giving the full hypothetical \$100: "I believe the work of such organizations has an urgency that affects many refugees and their families. If they can function without being affected by the harsh laws against non-profits, I would want them to have as much help as possible to do their humanitarian work."[^1097165] Thus, legal crackdowns can increase individual donor urgency to support an organization, and this heuristic is potentially processed differently compared to other organization-level heuristics or information from social networks.

In contrast, crackdowns have no substantial effect on the amount donated to privately funded humanitarian NGOs (`r filter(amount, title == "Humanitarian assistance | Private")$p_neq0`). No respondent in these conditions mentioned funding sources in their justification, and many explained that they would not give solely because of the crackdown and that they "would keep all \$100 because it would not even go to helping humanitarian efforts due to the new laws."[^8995923] It is possible that donors see government-funded humanitarian NGOs as more legitimate and more capable of handling difficult legal restrictions than privately funded organizations. Donors might assume that an NGO that receives government funding would have government support when facing legal challenges, while an NGO funded by small donors would be less able to confront such challenges. Further research is needed to probe this trend.

# Discussion

While the results from the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are a few caveats to consider when interpreting these results, generalizing them to other contexts, and applying them to nonprofit fundraising. First, the experiment measures preferences (willingness to donate) rather than behavior (actual donations). While the former may not always neatly translate into the latter, there is evidence that individuals’ willingness to donate (as measured in the experiment) translates well into behavior. The wording of the survey experiment uses an informational frame, where the focus is on educating respondents, rather simply presenting them with facts. Prior research has shown that informational frames—as opposed to the use of personal frames, which draw the audience’s attention to the plight of an individual, or motivational frames, which are meant to motivate individuals to act by creating feels of agency and efficacy—are important in generating donations [@McEntireLeibyKrain:2015].[^frametypes]

Moreover, prior research has found that informational frames are a necessary precondition for encouraging action in human rights nonprofit campaigns, and individuals are more likely to donate and undertake other actions when these informational frames are combined with personal frames [@McEntireLeibyKrain:2015].[^personalframes] This suggests that if governments attempt to repress international human rights nonprofits, then these organizations can potentially increase their resource flows by using a combination of these two frames. The survey instrument used above shows substantive increase in donors' willingness to donate to human rights groups facing crackdown, especially those that are privately-funded, and this result emerges from the use of just informational frames. If combined with personal frames, research indicates these donations may further increase.

It is also worth noting that there is evidence, at least domestically, that restrictions on nonprofit activities spur changes in private donors’ actual giving patterns (and not merely their preferences). For instance, following the 2016 U.S. presidential election, after Texas pledged to remove Planned Parenthood from its Medicaid program and a number of other states proposed new restrictions on their work, donations to the organization increased by over 40 times the usual amount [@citeme-as-well].

# Conclusion

Dozens of countries have sought to restrict NGO funding and programming over the past few decades. Between the lack of a local philanthropic culture, an increasingly authoritarian approach that restricts space for domestic philanthropy, and the withdrawal of official governmental aid, many international NGOs have turned to foreign private donors as an additional funding source. Since seeking out perfect information about a nonprofit and its activities—especially international programming—is costly and time-consuming, donors use heuristics when deciding which organizations to support. Existing research has shown that donors take cues and heuristics from their social networks and from organizational characteristics like GuideStar ratings. In this paper, we argue that other informational cues can influence donor preferences. We find that the domestic political environments of NGO host countries can indeed serve as one such heuristic. Our results show that while crackdowns do not consistently directly influence individual preferences of donation on their own, respondents indicated willingness to donate `r percent_0(filter(amount, title == "Crackdown")$pct_change)` more to legally restricted NGOs. This effect persists when mediated by other organizational heuristics, with respondents giving `r percent_0(filter(amount, title == "Humanitarian assistance")$pct_change)` more to humanitarian NGOs, and `r percent_0(filter(amount, title == "Human rights | Private")$pct_change)` more to privately funded human rights NGOs. Anti-NGO crackdowns can thus serve as a signal of NGO effectiveness.

This article should be viewed as an exploratory study. While the survey experiment allows us to hold the organization constant and test for specific hypotheses, there are limits to its external validity. Our sample captures people who are relatively more likely to donate to charities and thus does not reflect the general U.S. population. Our findings do not apply to the total population—we still know very little about how anti-NGO crackdowns change perceptions of nonprofits in general. Future research should look to testing the generalizability of our findings to individual donors in other countries, as well as determining the extent to which these preferences translate into behavior. For instance, the nonprofit sector in the U.S. relies much more heavily on government funding than in Europe, where large organizations like Amnesty International and Médecins Sans Frontières receive very little support from their home governments [@Stroup:2012], which means that donors in these countries will have different preferences when choosing to donate to government- and privately-funded organizations. Future research should also look more narrowly at donors who regularly give to international causes, as these individuals are likely more aware of poltiical trends in the countries where their preferred organizations work.

Because our sample reflects those who are already willing to donate to charities, these results have important implications for nonprofit fundraising. NGOs may benefit from publicizing when they are targets of government crackdown and including that information in their framing appeals. While this may not be enough to sway the average U.S. donor, our results show that communicating this information can convince already-likely donors to donate more to besieged groups. Further, if NGOs signal to donors that they receive a majority of their funding from nongovernmental sources, they may be able to convince average donors that their contributions might help the NGO resist the crackdown. Additionally, since they run the risk of receiving lower support under legal crackdown, human rights organizations that face legal difficulty abroad should communicate why they are under crackdown to mitigate donor mistrust.

[^deviations]: First, in the text of our preregistration, we inadvertently only specified non-nested versions of our hypotheses, but in our analysis plan, we describe the nested hypotheses. Second, in our analysis plan we stated that we would test for statisitcal differences using regression models with increasing numbers of interaction terms. This became unwiedly when interpreting results, though, and we instead simplified the analysis by estimating group means directly. The estimates are identical.

[^consort]: The CONSORT diagram in the online appendix shows the balance of respondents across different vignette combinations.

[^collapse]: An ordered probit model in the online appendix shows that results are consistent when using all five possible responses.

[^mturk]: Samples from Mechanical Turk are not nationally representative. However, because we targeted our study at people who are willing to donate online, we are not overly concerned with these discrepancies. Moreover, a growing body of research comparing Mechanical Turk with other sampling services finds it generally commensurable—see the online appendix for a lengthier discussion about this research.

[^targeting]: While anti-NGO legislation can theoretically target all NGOs within a state's borders, in practice these laws are enforced more selectively against contentious groups that pose threats to states [see @Chaudhry:2016; @Heiss:2017].

[^6216778]: Response 6216778.

[^13685]: Response 13685.

[^3025921]: Response 3025921.

[^6486084]: Response 6486084.

[^9509936]: Response 9509936.

[^1097165]: Response 1097165.

[^8995923]: Response 8995923.

[^8425616]: Response 8425616.

[^internationalaffairs]: Giving USA defines the sector of international affairs to include any nonprofit organization working in international development, international relief services, disaster relief, international human rights, international peace and security, foreign policy research and analysis, and international exchange programs. This classification is based upon the National Taxonomy of Exempt Entities Core which codes groups based upon their field of interest.

[^exceptions]: For exceptions, see @Micklewright:2009 on UK, @Rajan:2009 on Canada, @wiepking2010 on Netherlands, @CasaleBaumann:2015 on the USA.

[^highnetworth]: High-net worth households are defined as having greater than $200,000 in annual household income and/or with more than $1,000,000 in assets, excluding primary residence.

[^framing]: Framing is the process through which actors present information to influence perceptions or behavior. Previous research has shown that which frames are employed can have a significant effect on individuals' attitudes and behaviors [@BenfordSnow:2000; @ChongDruckman:2007].

[^frametypes]: For more on motivational and personal frames, see @BenfordSnow:2000 and @Gamson:1995.

[^personalframes]: Personal frames seek to humanize victims and entice action by invoking an emotional reaction. Research has found that people are more likely to be generous when donating to a human rights campaign with a single victim, as opposed to a statistical group of victims [@Slovic:2007].

[^NGO]: Nonprofit organizations (which we interchangeably use with NGOs) are organizations for a public or mutual benefit other than generating profit for the owners or investors [@Salamon:1999]. NGOs can be either domestic (i.e. an NGO that operates entirely in a single state), or international, more commonly known as INGOs. INGOs or international nonprofits, "operate independently of governments, are composed of members from two or more countries, and are organized to advance their members' international goals and provide services to citizens of other states through routine transactions with states, private actors and international institutions" [@Tarrow:2001]. Unlike domestic NGOs, INGOs often work in countries outside of their main country of origin or headquarters.